TRPV4 overactivation enhances cellular contractility and drives ocular hypertension in TGFβ2 overexpressing eyes

Version 1

Abstract

Abstract

The risk for developing primary open-angle glaucoma (POAG) correlates with the magnitude of ocular hypertension (OHT) and the concentration of transforming growth factor-β2 (TGFβ2) in the aqueous humor. Effective treatment of POAG requires detailed understanding of interaction between pressure sensing mechanisms in the trabecular meshwork (TM) and biochemical risk factors. Here, we employed molecular, optical, electrophysiological and tonometric strategies to establish the role of TGFβ2 in transcription and functional expression of mechanosensitive channel isoforms alongside studies of TM contractility in biomimetic hydrogels, and intraocular pressure (IOP) regulation in a mouse model of TGFβ2 -induced OHT. TGFβ2 upregulated expression of TRPV4 and PIEZO1 transcripts and time-dependently augmented functional TRPV4 activation. TRPV4 activation induced TM contractility whereas pharmacological inhibition suppressed TGFβ2-induced hypercontractility and abrogated OHT in eyes overexpressing TGFβ2. Trpv4-deficient mice resisted TGFβ2-driven increases in IOP. Nocturnal OHT was not additive to TGFβ- evoked OHT. Our study establishes the fundamental role of TGFβ as a modulator of mechanosensing in nonexcitable cells, identifies TRPV4 channel as the final common mechanism for TM contractility and circadian and pathological OHT and offers insights future treatments that can lower IOP in the sizeable cohort of hypertensive glaucoma patients that resist current treatments.

Introduction

Primary open-angle glaucoma (POAG) is an irreversible blinding disease afflicting ∼3.5% of the global population (1). Its incidence and severity are proportional to the amplitude and duration of ocular hypertension (OHT) (2, 3), which correlates with retinal ganglion cell dysfunction, neuroinflammation, and oxidative stress (4, 5). Biomechanical factors, glucocorticoids, and the cytokine transforming growth factor-β2 (TGFβ2) contribute to POAG by compromising the funneling of aqueous humor (AH) from the trabecular meshwork (TM) into the canal of Schlemm (SC). IOP elevations increase the contractility of juxtacanalicular TM (JCT), a circumocular tissue formed by extracellular matrix (ECM) beams populated by mechanosensitive cells smooth muscle-like cells, by increasing its resistance to the flow of aqueous humor (AH). The molecular mechanism that links TM pressure sensing to the contractile response is not known but is likely to underpin the uts sensitivity to compressive, tensile, osmotic, shear and traction forces to regulate the expression of hundreds TM genes and secretion of dozens of ECM proteins (6–12).

The increase in trabecular outflow resistance induced by mechanical stress, glucocorticoids, and TGFβ2 consists of a dynamic component that can be reversed by cytoskeletal and Rho signaling blockers, and a chronic component, reflecting transdifferentiation of TM cells into fibrotic and contractile myofibroblasts (16–18). One of the clearest examples of fibrotic remodeling in glaucoma pertains to TGFβ2 signaling: (i) TM cells derived from POAG patients secrete more active TGFβ2 compared to cells isolated from healthy donors (19), (ii) the likelihood of contracting POAG is proportional to [TGFβ2]AH (20–22), and (iii) ocular overexpression of TGFβ2 is sufficient to induce OHT (23, 24), presumably due to overexpression of ECM proteins and increased cellular contractility (25, 26). The cognate TGFβ1 isoform similarly induces fibrotic remodeling in fibroblasts, epithelial, and endothelial cells in the heart, kidney, skin, and lung (27–30). TGFβ overexpression is thus a causal determinant of OHT that may reflect a universal fibrotic program that, however, cannot be disambiguated from the biomechanical environment. TGFβ release is activated by tissue contractility and tension (31, 32), TGFβ activity correlates with the distribution of mechanical stress (33), and mechanical stress may induce TGFβ - dependent epithelial-mesenchymal transition (EMT; 34, 35).

Our understanding of TM mechanotransduction and its role in IOP homeostasis remains rudimentary despite its overarching clinical relevance. Strain and shear were proposed to stimulate the TM primary cilium, integrins and TRPV4, Piezo1 and TREK-1 channels (36–39) but it is not clear whether mechanosensation regulates TM contractility or is itself impacted by POAG inducers (e.g., TGFβ2 or glucocorticoids) nor is it known how it relates to chronic fibrosis. TRPV4 (Transient Receptor Potential Vanilloid isoform 4), a tetrameric channel with PCa/PNa ∼ 10 (40), is strongly expressed in rodent and human TM (36, 51) where it carries the principal component of the pressure-activated transmembrane current and stretch-evoked [Ca2+]i elevations together with responsiveness to shear and swelling (7, 10, 37, 39, 41, 42). Pharmacological inhibition of the channel and deletion of the TRPV4 gene modulate pressure gradients in the brain, kidney, lung, and bladder (46–50) and mutations in the TRPV4 sequence underpin sensorimotor neuropathies, skeletal dysplasias, retinal degeneration and ocular dysfunction (43–45) while the function of TRPV4 channels in ocular hypertension remains under debate. TRPV4 activity has been implicated in IOP lowering and elevation, respectively and linked to a diverse array of effector mechanisms that include eNOS and RhoA activation, phospholipid-cholesterol-caveolin regulation, modulation of cell-ECM contacts, primary cilia mechanosensing, polyunsaturated fatty acid release, and Piezo1 signaling (7, 37, 41, 52–55). This invites testable predictions regarding TRPV4 involvement in ocular function and POAG. If TRPV4 maintains steady-state normotension, opposes the profibrotic effects of TGFb2 and promotes outflow via eNOS-dependent TM relaxation (7, 52), abrogation of its activation should elevate IOP. Conversely, if TRPV4 promotes OHT (37), its inhibition of deletion of the TRPV4 gene should lower IOP.

In this study, we demonstrated novel functions for the TRPV4 channel in homeostatic and pathological IOP regulation by uncovering the mechanisms through which reciprocal TRPV4-TGFb2 interactions maintain the vicious cycle between mechanical stressors and TM contractility that underlies OHT. We found that inhibition and deletion of TRPV4 lower IOP in TGFβ2 overexpression-induced and circadian models of OHT and suppress TM contractility in TGFβ2-treated biomimetic hydrogels. The cytokine promoted upregulation of EMT-associated genes alongside increased transcription and trafficking of TRPV4, which may have increased the sensitivity of TM cells to innocuous mechanical stimuli. While physiological (nocturnal) and pathological (cytokine-induced) OHT modes both required TRPV4 activity they were not additive, indicating the involvement of a final common pathway.

Collectively, we identify TRPV4 as a fulcrum of TGFβ2 -induced TM contractility and IOP regulation and a candidate target for glaucoma therapy.

Results

TGFβ2 drives overexpression of genes that encode fibrotic markers and mechanosensitive ion channels

Human TM cells respond to TGFβ2 with increased biosynthesis, deposition and degradation of ECM, altered autophagy, upregulation of F-actin stress fibers, a-smooth muscle actin (aSMA) (19, 25, 26, 56, 57), but it is unclear whether cells undergoing TGFβ2-induced fibrotic remodeling also exhibit altered capacity for sensing and transduction of mechanical stimuli. We profiled genes that encode known TM mechanochannels together with a selection of key cytoskeletal, ECM, and fibrotic markers in primary TM cells (pTM) isolated from 3-7 donors without history of visual dysfunction (Figure 1A-C). Five-day exposure of pTM cells to a physiological concentration of TGFβ2 (1 ng/mL) increased the expression of EMT-promoting transcription factor SNAI1 (SNAIL1, P = 0.0094) and fibronectin (FN1, P = 0.0263), while expression of connective tissue growth factor 2 (CCN2, alternatively CTGF) was elevated in 5/5 pTM cell strains without reaching significance (P = 0.0909). Expression of fibroblast-specific protein 1 (FSP1, a calcium-binding fibroblast marker), yes-associated protein 1 (YAP1, a stiffness induced hippo- pathway transcription factor) and ACTA2 (αSMA, associated with cell contractility) was not consistently impacted by TGFβ2 while transcription of myocilin (MYOC) decreased across 4/4 pTM strains (P = 0.0055) (Figure 1B). Indicative of feedback inhibition (58), TGFβ2-treatment downregulated transcript levels of transforming growth factor beta receptor 2 (TGFBR2, P = 0.0219) and upregulated expression of autoinhibitory SMAD family protein 7 (SMAD7, P = 0.0461) without affecting SMAD2 or SMAD3 expression. TGFβ2 thus promotes selective upregulation of ECM and fibrosis-related genes together with cell dedifferentiation and activation of autoregulatory SMAD mechanisms.

TGFβ2 induces a fibrotic phenotype in pTM cells and increases expression and membrane insertion of the TRPV4 channel.

(A-B) Five-day TGFβ2 treatment (1ng/mL) significantly altered expression of TGFβ pathway effectors, cytoskeletal machinery, and canonical fibrotic markers. (C) TGFβ2 treatment significantly increased TRPV4 and PIEZO1 expression, but not TREK1 and TRPC1 expression. Mean ± SEM shown. N = 4 - 8 experiments, each gene tested in 3-7 different pTM strains (See Table 1). Two-tailed one sample t-test of TGFβ2-induced gene expression levels as a percent of control samples. (D) Isolation of membrane proteins from two separate pooled pTM samples suggests TGFβ2 treatment drives increased TRPV4 membrane insertion. N = 2 independent pooled samples, 3 pTM strains were pooled per sample. * P < 0.05, ** P < 0.01.

Analysis of genes encoding mechanosensitive channels implicated in outflow modulation (36, 39, 59, 60) showed a 102.5% increase in expression of TRPV4 (P = 0.0193) and 78.9% increase in PIEZO1 expression (P = 0.0114) across 8 replicates including 7 distinct pTM strains. (Figure 1C). Conversely, TGFβ2 exposure did not affect TRPC1 gene expression (P = 0.261) and had variable, strain-dependent effects on transcript levels of KCNK2 (P = 0.293, encoding the TREK-1 channel). Thus, TGFβ2 promotes selective transcriptional upregulation of genes that encode a subset of mechanosensitive proteins alongside fibrotic upregulation and cell dedifferentiation.

TGFβ2 exposure time-dependently augments TRPV4-mediated current and Δ[Ca2+]i

To assess the functional relevance of TGFβ2-dependent transcriptional upregulation we determined the membrane expression and functional activation of TRPV4, which mediates the pressure- activated current and calcium signaling, regulates cytoskeletal dynamics and modulates conventional outflow resistance in vitro (37, 41). Western blots showed that TGFβ2 exposure produces an increase in levels of membrane-bound TRPV4 protein (Figure 1D) in two grouped pTM membrane protein samples. While low amounts of TRPV4 were visible in the membrane fractions in control samples, TGFβ2 treatment produced an increase in the higher weight TRPV4 band, suggesting there could be isoform- specific TGFβ2-induced responses and increased TRPV4 translation leading to elevated TRPV4 trafficking, membrane insertion and/or lipid raft interaction (52).

Functional expression was assessed by tracking [Ca2+]i changes in cells exposed to the selective TRPV4 agonist GSK1016790A (GSK101, 10 nM) using Fura2-AM ratiometric Ca2+ dye, with TGFβ2-treated and control cells tested on the same day. All pTM strains responded to GSK101 with robust [Ca2+]I increases which reached peak within 5 min before the majority of responding cells gradually decreased to a steady plateau (Figure 2C). TGFβ2-treated cells exhibited a remarkable potentiation of GSK101-evoked [Ca2+]i responses compared to control cells, with 5/5 cell strains showing an increase in the Δpeak/baseline F340/F380 response equivalent to 258.4% ± 61.7% of the control response in (P = 0.0046) (Figure 2A-B). The fraction of GSK101 responders and the overall time course of responses between groups were not significantly different, indicating that TRPV4 potentiation primarily affects TRPV4- expressing cells. Thus, TGFβ2 treatment promotes TRPV4 expression and functional activity.

TRPV4-mediated Ca2+ influx is potentiated by five-day TGFβ2 treatment.

(A) Five-day TGFβ2 treatment (1 ng/mL) increased TRPV4 agonist-induced (GSK101, 10 nM) Ca2+ influx in pTM cells compared to serum-free media alone treated cells tested on the same day (N = 5 pTM strains, n = 3 - 5 slides/condition/day, individual data points over mean ± SEM). Two-tailed one sample t- test of TGFβ2-treated cell average GSK101 response as a percent of control samples from the same pTM strain on the same day. (B) Violin plots showing the distribution of GSK101-induced Ca2+ responses for each pTM strain tested in A. Thick dashed line indicates mean, while light dashed line indicates quartiles. (C) Representative traces showing TRPV4 agonist-induced Ca2+ influx (seen as an increase in F340/F380) in pTM (mean ± SEM of 4 representative cells/ group), alongside example Fura-2-loaded pTM cells before (i), during (ii), and after (iii) GSK101 application. Scale bar = 50 µm. ** P < 0.01

To gain insight into the time- and dose-dependence of TGFβ2-dependent TRPV4 signaling modulation pTM cells were treated for 24 hours, at 1 ng/mL and 5 ng/mL concentrations of TGFβ2. GSK101-stimulated Ca2+ influx was not significantly increased by 24-hour TGFβ2 treatment at 1 ng/mL (Δpeak/baseline F340/F380 = 117.0% ± 23.6% of control) or 5 ng/mL (Δpeak/baseline F340/F380 = 133.6% ± 34.5% of control) (Figure 3; SI Appendix, Figure S1); additionally, the potentiation of both was significantly lower relative to the five-day 1 ng/mL TGFβ2 treatment (P < 0.0011; Figure 3A). GSK101 evoked a moderately outwardly rectifying nonselective current (IGSK-Ibaseline) with reversal potential at ∼0 mV (Fig 3C). While its amplitude was variable, mean current density consistently increased in cells treated for 1 day with TGFβ2 (n = 10; 5 ng/mL) relative to the control group (n = 11). The potentiating effect of TGFβ2 on TRPV4 activity thus appears to be time-dependent but is significant after chronic exposure to relatively low-dose TGFβ2.

TGFΒ2-induced TRPV4 potentiation is not seen at a shorter time period, regardless of treatment strength.

(A) TGFβ2 treatments for 24 hours at 1 ng/mL (N = 6 pTM strains, n = 3 - 5 slides/condition/day) or 5 ng/mL (N = 5 pTM strains, n = 3 - 5 slides/condition/day) did not show potentiation of GSK101-evoked TRPV4 Ca2+ influx (SI Appendix, Figure S1) and were significantly lower than cells treated with TGFβ2 for 5d at 1ng/mL (5d TGFβ2 results from Figure 2A). Individual data points over mean ± SEM. One-way ANOVA with Tukey’s multiple comparisons test, statistics for individual 1d treatment groups compared to control groups shown in Figure S1. (B) Representative traces for GSK101 response following 24-hour TGFβ2 treatment, traces show mean ± SEM of 3-4 cells. (C) Average current density in response to GSK101 (24-hour control: n = 11 cells, 24-hour TGFβ2: n=10 cells) shows generally increased current in TGFβ2-treated cells. Data shows mean ± SEM (D - E) Violin plots of individual cell strains shown in A. Thick dashed line indicates mean, while light dashed line indicates quartiles. ** P < 0.01, *** P < 0.001

TGFΒ2-Induced TM contractility requires TRPV4 activation

The IOP-lowering effectiveness of Rho kinase inhibitors and latrunculins (57, 61–63) indicates that sustained increases in outflow resistance require tonic actin polymerization and contractility. TGFβ2 drives the TM myofibroblast contractile response (57) while the role of mechanosensation remains unknown. To ascertain whether TRPV4 upregulation (Figures 1-2) contributes to the contractile response, we seeded pTM cells into high-compliance Type I collagen hydrogels (57) (Figure 4, SI Appendix, Figure S2). Hydrogels that were incubated with TGFβ2 (5 ng/mL) showed profound increases (P < 0.0003) in the rate and the magnitude of contraction at all time points (Figure 4, SI Appendix Figure S2).

Effects of TRPV4 inhibition/activation on TGFβ2-induced contraction of TM cells.

(A) Representative longitudinal 24-well plate scans of collagen type I hydrogels seeded with pTM subjected to the different treatments (dashed lines outline size of contracted constructs). (B) Longitudinal quantification of hydrogel construct size compared to the control group at the 0 minute time point. (C) Detailed comparisons between groups at each experimental time point. n = 6 hydrogels/group. One-way ANOVA with Tukey’s multiple comparisons test, data in (B,D) shows individual data points over mean ± S.D. One pTM strain shown: TGFβ2-induced contractility induction, HC-06-mediated rescue of hypercontractility, and GSK101-induced transient (15 min) contraction were consistent across (3/3) pTM strains tested (Figure S2). ** P < 0.01, *** P < 0.001, **** P < 0.0001.

Simultaneously, treatment with the TRPV4 antagonist HC-067047 (HC-06, 5 µM) significantly reduced the extent of TGFβ2-induced TM contractility (P < 0.0001). To determine whether TRPV4 activation is sufficient to induce the contractile response, the antagonist was washed out and hydrogels supplemented with GSK101 (25 nM). 15 minutes post-treatment, the constructs responded to the agonist with transient contraction (Figure 4C; SI Appendix, Figure S2, P < 0.01), with a time course mirroring GSK101- induced [Ca2+]i elevations (Figures 2-3). The effects of TRPV4 inhibition and activation were consistent across all pTM strains tested (N = 3 pTM strains). TRPV4-mediated Ca22+ influx is therefore sufficient to induce TM contractility and necessary for pTM hypercontractility induced by TGFβ2.

TRPV4 activation is required to maintain TGFβ2-induced OHT

To test whether TRPV4 contributes to TGFβ2 induced ocular hypertension (OHT) in vivo, we utilized the lentiviral TGFβ2 overexpression model developed by Patil et al. (23). Adult C57BL/6J mice (N = 5) were intravitreally injected with lentivirus overexpressing constitutively active human TGFβ2 (LV-TGFβ2). LV-TGFβ2-injected eyes, but not the contralateral eyes injected with a lentivirus containing a scrambled transgene (LV-Ctrl), exhibited significant IOP elevations one-week post-transduction (Figure 5A, Week 2, ΔTGF-Ctrl = 4.0 mm Hg, P = 0.0143). By 2 weeks post-transfection, IOP in LV-TGFβ2 eyes reached 19.9 ± 4.7 mm Hg whereas IOP in LV-Control eyes remained at control levels (14.0 ± 1.2 mm Hg), with ΔTGF-Ctrl = 5.9 mm Hg (P = 0.0002). IOP remained elevated throughout the 4 weeks after the injection (Week 5, ΔTGF-Ctrl = 4.9 mm Hg, P = 0.0008). HC-06 (100 µM) microinjection into the anterior chamber of LV-TGFβ2 and LV-Ctrl eyes lowered IOP in LV-TGFβ2 eyes to 12.2 ± 1.7 mm Hg after 24 hours (ΔpostHC-preHC = -5.8 mm Hg) with no difference observed in IOP from LV-Ctrl eyes (12.6 ± 1.9 mm Hg, Δpost-HC-pre-HC = -0.3 mm Hg). LV-Ctrl eyes remained close to pre-injection levels post-HC-06 treatment (Figure 5A-B). IOP in LV-TGFβ2 eyes returned to hypertensive levels by 1-week post-HC-06 injection (Week 6-7, ΔTGF-Ctrl = 3.9 mm Hg, P = 0.0201). To determine the effect of the bolus injection alone, LV-TGFβ2 and LV-Ctrl eyes were reinjected with PBS 2 weeks after re-establishing the OHT baseline. The sham injection transiently reduced IOP in LV-TGFβ2 (ΔpostPBS-prePBS= -4.5 mm Hg) and LV- Ctrl (ΔpostPBSpre-PBS= -1.2 mm Hg) eyes; however, LV-TGFβ2 eyes returned to hypertensive levels by 48 hours post-injection (ΔTGF-Ctrl =3.6 mm Hg, p=0.0465) and to pre-injection levels after 72 hours (ΔTGF-Ctrl =5.4 mm Hg, p=0.0002). Bolus injection was less effective than HC-06 at all time points 24 hours post- injection (Week 8-9, Figure 5B). These data indicate that selective pharmacological inhibition of TRPV4 effectively and reversibly blocks TGFβ2-induced OHT.

TRPV4 activation is necessary to maintain LV-TGFβ2-induced ocular hypertension.

(A) Intravitreal injection of LV-TGFβ2 (week 1), but not LV-Control, elevates IOP in WT mice (N = 5 eyes/group) as early as one-week post-injection. Injection of TRPV4 antagonist HC-06, but not PBS, produced multiday IOP reduction in LV-TGFβ2 treated eyes. HC-06 and PBS injections did not affect IOP in LV-Control injected eyes. Two-way ANOVA with Bonferroni post-hoc analysis (B) Direct comparison of the results of PBS and HC-06 injections in the eyes shown in A. Two-way ANOVA with Bonferroni post-hoc analysis (C) Intravitreal injection of LV-TGFβ2 in Trpv4-/- mice (N = 6 eyes/group) resulted in only mild OHT; plotted against WT eyes at matching timepoints (3 WT cohorts including the 5 WT eyes shown in A-B, N = 8-15 eyes/group). (D) Statistical comparison of the IOP values shown in C. The IOP in LV-TGFβ2 WT eyes was significantly elevated compared to the LV-TGFβ2 Trpv4-/- eyes from 2 weeks post-injection. LV-Control injected eyes in WT or Trpv4-/- eyes remain close to the baseline value and are not significantly different. Two-way ANOVA with Bonferroni post-hoc analysis. (A, C) shows mean ± SEM. Data in (B, D) shows individual data points over mean ± SEM, * P < 0.05, ** P < 0.01, *** P < 0.001, **** P < 0.0001

To further evaluate the TRPV4-dependence of TGFβ-induced OHT we took advantage of mice with global Trpv4 knockdown (64–66). Trpv4-/- mice (N = 6) were injected with LV-TGFβ2 and LV-Ctrl vectors in contralateral eyes (Figure 5C). Additionally, two littermate control mice injected alongside the Trpv4-/- animals were added to previously collected WT LV-injected cohorts measured at the same timepoints (N = 8-15, Figure 5C). Pre-LV injection, IOP levels in Trpv4-/- animals were comparable to the WT cohort, indicating that TRPV4 activity does not regulate normotension. Similarly, IOP in LV-Ctrl- injected eyes was not significantly different between WT and Trpv4-/- animals at any point in the experiment (peak ΔCtrlKO-CtrlWT = -1.2 mm Hg, SI Appendix, Figure 5D). By two weeks post-injection (Week 3), LV-TGFβ2-treated Trpv4-/- eyes exhibited significantly lower IOP compared to the LV-TGFβ2 WT cohort (ΔTGFKO-TGFWT = -3.1 mm Hg, P =0.0009, Figure 5C). While LV-TGFβ2 injected Trpv4-/-eyes exhibited mild OHT, the effect was significantly reduced compared to WT eyes and IOP decreased by two weeks post-injection (Figure 5C-D).

TGFβ2-induced and nocturnal OHT are non-additive but require TRPV4

Mammalian IOP is modulated by the circadian rhythm, with levels elevated at night and nocturnal IOP fluctuations implicated in POAG (7, 55, 67). We measured nocturnal (9:00-10:00 PM) IOP in LV- TGFβ2 and LV-Ctrl WT eyes (N=4) to determine whether nocturnal OHT is additive to elevation observed during the daytime (12:00-2:00 PM, Figure 6A). LV-TGFβ2 injected eyes showed significant IOP elevation compared to LV-Ctrl eyes during daytime measurements (Diurnal ΔTGFβ-Ctrl = 7.9 mm Hg, P < 0.0001) but the difference vanished at night (Nocturnal ΔTGFβ-Ctrl = 0.2 mm Hg), indicating that TGFβ2- induced OHT is not additive to the circadian OHT. To determine whether physiological (nocturnal) OHT requires TRPV4 we microinjected the eyes of two animals with PBS, and two with HC-06. When IOP stably recovered after the first treatment, the treatment groups were switched. PBS injection did not affect IOP in LV-Ctrl or LV-TGFβ2 eyes at day or night (Figure 6B-C) except for a single LV-TGFβ2 eye exhibiting abnormally high nocturnal IOP (37 mm Hg) at the four-day timepoint. Conversely, HC-06 injection blocked LV-TGFβ2-induced IOP during the day (P < 0.001) and significantly lowered IOP ∼5 mm Hg in both LV-Ctrl and LV-TGFβ2 eyes at night (P < 0.01). These data indicate that i) TRPV4 activation is necessary for OHT in the TGFβ2 overexpression mouse model (Figures 5-6) and the circadian IOP elevations ii) TGFβ2 -evoked OHT does not affect nocturnal IOP elevation in mice, and iii) TRPV4 inhibition does not disrupt the mechanisms that maintain daytime normotensive IOP (Figures 5- 6).

TRPV4 inhibition inhibits nocturnal IOP elevation in control and TGFβ2 overexpressing eyes.

(A) ∼2 months post-LV injection daytime (12-2:00 P.M) and nocturnal (9-10:00 P.M.) IOP compared in WT mice (N = 4 eyes/group) before drug treatment. LV-TGFβ2 eyes were elevated at daytime, but nocturnal OHT was not significantly different between LV-Ctrl and LV-TGFβ2 eyes. One- way ANOVA with Tukey’s multiple comparisons test. (B-C) PBS-injected eyes did not exhibit changes in daytime or nighttime intraocular pressure; however, HC-06 injection reduced TGFβ2-induced IOP elevations during the day and LV-Ctrl and LV-TGFβ2 nocturnal IOPs (N = 4 Eyes/Group); Two-way ANOVA with Bonferroni post-hoc analysis. Figures show datapoints over mean ± SEM, *P < 0.05, ** P < 0.01, *** P < 0.001, **** P < 0.0001

Discussion

The mechanistic framework developed in this study unifies key biochemical and biomechanical risk factors of POAG to point towards an alternative approach to mitigate fibrotic and functional dysfunction in eyes experiencing OHT. Specifically, we show that TGFβ2 drives overexpression and excessive activation of TRPV4, a Ca2+-permeable channel with diverse mechanosensing functions that include mediating the principal component of the pressure-activated transmembrane current roles in TM cells and fibrotic remodeling across the body (39, 68, 69). Our central observation - that tonic TRPV4 activity is obligatory to maintain TM contractility and OHT induced by angle occlusion and TGFβ2 – identifies a potential molecular linchpin for increased resistance of the JCT TM to AH outflow.

Considering that current glaucoma treatments target secondary outflow mechanisms or are associated with side effects (such as hyperemia) (70, 71), the IOP lowering effected by TRPV4 inhibition and gene knockdown suggests a novel target within the primary outflow pathway that can be engaged without compromising the structural integrity or function of the anterior eye.

Glaucoma is a multifactorial disease with etiology that reflects the convergence of risk factors that include IOP and TGFβ2: the likelihood of POAG correlates with the amplitude of IOP and [TGFβ2]AH (22, 72), and chronic increases in either [TGFβ2]i or IOP promote fibrotic remodeling of the TM/SC and augment the AH flow resistance of the conventional pathway (17, 24, 25). TGFβ2-induced facility suppression has been historically attributed to changes in composition, crosslinking and amount of ECM (25, 26, 73, 74), activation of Hippo signaling and Rho kinase- (Rho/ROCK) mediated contractility (19, 57) and altered expression of genes encoding mitogen-activated protein kinase (MAPK), immune response, oxidative stress, and/or ECM pathways (75–77). Our discovery that TGFβ2 impacts the expression and function of TM mechanosensors and vice versa, that TRPV4 is required for TGFβ2- induced contractility, coalesces two key modifiable risk factors (TGFβ2 and pressure) at the level of TRPV4 signaling. Specifically, our data embed TGFβ2 and TRPV4 signaling within reciprocal feedback loops: TGFβ2 (i) induced time-dependent upregulation of TRPV4 mRNA and protein and amplified TRPV4-mediated calcium signaling, while (ii) TRPV4 was required to mediate TGFβ2 -induced contractility and maintain chronic OHT in TGFβ2-treated mouse eyes. Microinjection of the selective antagonist HC-06 accordingly reduced IOP in LV- TGFβ2-treated eyes to baseline with hypotension persisting for ∼4 days and reversing to pre-injection OHT by day 7. The TRPV4-dependence of TGFβ2- induced OHT and contractility was validated in vivo using Trpv4-/- eyes and in vitro in 3D hydrogel constructs. The differential effectiveness of IOP lowering induced by gene knockdown (∼50% reduction in OHT) and pharmacological inhibition (∼100% reduction in OHT) may reflect compensatory upregulation of cognate mechanosensory mechanisms in Trpv4-/- animals (55).

We’ve previously shown that TRPV4 channels in primary human TM cells are activated by physiological (5 – 25 mm Hg) pressure steps (39, 60) and (1 – 12%) strains (37, 41), to activate downstream outflow-relevant signaling mechanisms such as Rho kinase, F-actin, tyrosine phosphorylation of FAK, paxillin and vinculin, reorganization of membrane lipids, and ECM release (37, 41, 52). Here, we extend those observations to demonstrate that TRPV4 activation is required for TM contractility and ocular hypertension induced by TGFβ2 overexpression and circadian rhythmicity. The observation that TRPV4 activity underpins increased outflow resistance under physiological and pathological conditions resolves contradictory conclusions from prior investigations, which implicated TRPV4 signaling in ocular hypertension and hypotension, respectively (7, 36, 37, 39, 41, 52, 53, 55). TRPV4 has been proposed to lower IOP through phosphoinositide signaling in primary cilia (36), stimulate TM-resident endothelial nitric oxide synthase (eNOS) (7) and release of polyunsaturated fatty acids (PUFAs) (53) and/or activate downstream from Piezo1 mechanosensing (54). However, TRPV4- regulated Ca2+ influx in TM cells is unaffected by the ablation of primary cilia (37), eNOS expression in TM cells is vanishingly low (78–80), PUFAs such as arachidonic acid and EETs stimulate rather than inhibit, TRPV4 (37) and TRPV4 signaling in TM cells is unaffected by Piezo1 inhibitors and knockdown (39). Moreover, Piezo1 inhibition reduces outflow facility under in vitro and in vivo conditions (39, 81), indicating opposing homeostatic functions for Piezo1 vs. TRPV4 activation. The TRPV4-dependence of TM contractility (Figure 4) accords with reports that TRPV4 inhibition increases, and TRPV4 activation reduces outflow facility in biomimetic TM-populated scaffolds in the absence of ciliary body, Schlemm’s canal, and ciliary muscle (37). Induction of the contractile response by GSK101 and its inhibition of hypercontractility by HC-06 further suggest a model whereby TRPV4 pressure transduction drives Ca2+- and Rho-dependent hypercontractility and fibrosis via actin polymerization, myosin light chain phosphorylation, aSMA integration into stress fibers and reinforcement of focal ECM contacts (41, 82, 83) (Figure 7). TRPV4 channels in cells treated with TGFβ2 are likely to be constitutively active at incubator temperature, which coincides with peak TRPV4 thermoactivation (∼34 - 37°C) (84, 85). The residual contractility in HC-06-treated cells may reflect TGFβ2-mediated contributions from Piezo1, TRPC, and/or TREK-1 channels and/or intracellular Ca2+ release (37, 60, 86, 87). Reports from heart, lung, liver, skin and articular cartilage preparations similarly implicate TRPV4 in TGFβ1 -dependent fibrosis (68, 88–91) and bladder (92), heart (93, 94), and vascular (95) contractility, with conditional ablation of TRPV4 from smooth muscle cells shown to lower blood pressure (96, 97). TGFβ2-stimulated induction of FN1, SNAIL1, and CTGF transcripts (Figure 1) accords with RNA profiling studies which documented the cytokine’s key role in TM transdifferentiation towards the contractile myofibroblast state (25, 75, 98–103) whereas the decreased expression of TGFΒR2 and increased the abundance of SMAD7 mRNA indicate activation of autoinhibitory mechanisms in EMT -undergoing cells (104).

TGFβ2-TRPV4 interactions in TM remodeling and ocular hypertension.

Chronic exposure to TGFβ2 induces upregulation of functional TRPV4 channels alongside the autoinhibitory canonical modulator SMAD7. TRPV4-mediated Ca2+ influx, canonical, and non-canonical TGFβ2 signaling stimulate Rho/ROCK signaling, augment cytoskeletal contractility, and stimulate ECM release to increase the flow resistance of the conventional pathway. Increased contractility drives OHT, resulting in a feedforward vicious TRPV4-dependent circle loop that maintains ocular hypertension. Schematic made using Biorender.com.

Treatment of TM cells with TGFβ2 concentrations comparable to those found in POAG AH (0.2- 3.2 ng/ml) (20) produced 2-3-fold upregulation of TRPV4 transcripts, protein, and responsiveness to GSK101, with the time course of these effects mirroring facility reduction in human eyes treated with exogenous cytokine (105). A single exposure to 5 ng/ml TGFβ2 approximately doubled the amplitude of the GSK101-evoked current and reduced outward rectification of ITRPV4 (Figure 3). Difficulties with giga- ohm seal formation precluded ITRPV4 analyses at longer incubation times but we were able to obviate this limitation with imaging experiments, which revealed robust and reproducible time-dependent increases in the amplitude of agonist-induced Ca2+ signals across all 5 studied strains (Figure 2). The effects of TGFβ2 on ITRPV4, membrane protein levels and [Ca2+]GSK accord with increased expression of the TRPV4 gene, with precedents from other cell types (e.g., fibroblasts) suggesting the possibility of increased trafficking of TRPV4- PI3Kg complexes and/or β-arrestin 1-dependent ubiquitination (106, 107). The upregulation of TRPV4/Piezo1 transcription predicts that TGFβ2-exposed cells might exhibit exaggerated responsiveness to mechanical loading and mechanical hyperalgesia, as reported for chemotherapy (108), neuropathic pain (109, 110), cancer (111), and diabetic neuropathy (112).

The lack of additivity between TGFβ2-induced OHT and nocturnal OHT (Figure 6) suggests that control of physiological and pathological hypertension converges at the level of TRPV4-Rho signaling as the final common mechanism obligatory for OHT. This conjecture is supported by the observations that TRPV4 inhibitors, ROCK inhibitors and TM-specific expression of dominant negative scAAV2.dnRhoA constructs lower IOP, elevated through occlusion of the iridocorneal angle, TGFβ, glucocorticoids and the nocturnal cycle (55, 62, 113). Future studies of TRPV4 signaling will investigate the mechanisms that underlie the reversibility of circadian TRPV4 activation and how TGFβ2 impairs this physiological process. For example, the suprachiasmatic nucleus and the hypothalamus–pituitary–adrenal axis (114, 115) may modulate the TRPV4-Rho axis via nocturnal release of norepinephrine and melatonin (116, 117). It is worth noting that TRPV4 may directly bind to membrane proteins known to regulate conventional outflow, such as β1 integrins (118), caveolin-1 (52), and cytoskeletal proteins (actin, actin adaptor proteins, microtubules) (119).

Our study unifies biomechanical and biochemical paradigms of fibrotic and functional remodeling in glaucoma, expands the biological significance of TGFβ2 modulation by including increased actomyosin contractility in addition to fibrotic remodeling, thereby opening a new window into the mechanisms that subserve physiological vs. pathological OHT. We propose that TGFβ2 shifts the homeostatic normotensive setpoint maintained by steady-state TRPV4, Piezo1 and TREK-1 activation (39, 60, 81) through upregulation of TRPV4 expression, which increases the cells’ sensitivity to pressure and strain under normotensive conditions. In addition to hijacking the cells’ contractile apparatus, TGFβ2 overexposure induces fibrosis that may facilitate the pull of stress fibers on the increasingly “stiff” ECM (19) together with increased production and secretion of ECM. The absence of structural and functional visual phenotypes in TRPV4 KO mice (55, 66, 120) predicts that small-molecule TRPV4 antagonists might lower IOP, suppress fibrosis and protect retinal neurons without compromising homeostatic IOP regulation (70). The similarities between TRPV4 expression in mouse and human TM (37) and between outflow mechanisms in mice vs. humans (121, 122) suggest that TRPV4 targeting might be explored within the clinical context.

Methods

Animals

C57BL/6J mice were from JAX laboratories (Bar Harbor, ME), Trpv4-/- mice were a gift from Wolfgang Liedtke (Duke University) (64, 65). The animals were maintained in a pathogen-free facility with a 12- hour light/dark cycle and ad libitum access to food and water, at a temperature of ∼22-23°C. Mice were 2-6 months in age prior to LV-injection; data from both male and female sexed animals were included in this study.

Human TM Culture

De-identified postmortem eyes from donors with no history of glaucoma (pTM cells) were procured from Utah Lions Eye Bank with written informed consent of the donor’s families. TM cells were isolated from juxtacanalicular and corneoscleral regions as previously described (37, 39), in accordance with consensus characterization recommendations (123). pTM cells were cultured in Trabecular Meshwork Cell Medium (TMCM; Sciencell) in Collagen-I (Corning) coated culture flasks and glass coverslips at 37°C in a humidified atmosphere with 5% CO2. Fresh media was supplied every 2-3 days. Serum free (SF) media was mixed as needed by excluding fetal bovine serum (FBS, Sciencell) from the TMCM. A list of all pTM strains used is available in Table 1; all cells were used between passages 2-4. Cell lines were chosen based on availability at the time of experiments.

For contractility experiments pTM cells were isolated from healthy donor corneal rims discarded after transplant surgery, as previously described (19, 57, 124), and cultured according to established protocols (123, 125). Three pTM cell strains isolated from healthy donors and validated with dexamethasone- induced myocilin expression were used for contractility experiments. pTM cells were cultured in low- glucose Dulbecco’s Modified Eagle’s Medium (DMEM; Gibco; Thermo Fisher Scientific) containing 10% fetal bovine serum (FBS; Atlanta Biologicals) and 1% penicillin/streptomycin/glutamine (PSG; Gibco) and maintained at 37°C in a humidified atmosphere with 5% CO2. Fresh media was supplied every 2-3 days.

The experiments were conducted according to the tenets of the Declaration of Helsinki for the use of human tissue.

Reagents

The TRPV4 antagonist HC-067047 (HC-06) was purchased from Millipore-Sigma or Cayman Biotech and dissolved in DMSO at 20mM. The TRPV4 agonist GSK1016790A (GSK101; Cayman Biotech) was dissolved in DMSO at 1mM. Aliquots were diluted into working concentrations (10-25 nM, GSK101; 5- 100 µM, HC-06). Recombinant human TGFβ2 protein was ordered from R&D Systems and reconstituted in sterile 4 mM HCl with 0.1% BSA at 20 ug/mL.

Quantitative Real-Time PCR

Gene-specific primers were used to detect expression of target genes, as described (126). Total RNA was isolated using the Arcturus PicoPure RNA isolation kit (Thermofisher Scientific). cDNA was generated from total RNA using qScript XLT cDNA Supermix (Quanta Biosciences). SYBR Green based real-time PCR was performed with 2X GREEN Master Mix (Apex Bioresearch Products). Gapdh was used as an endogenous control to normalize fluorescence signals. Gene expression relative to GAPDH was measured using the comparative CT method (2-[ΔCT(gene)- ΔCT(GAPDH)]). All genes were assessed in 4-8 individual samples taken from 3-7 different pTM strains. The primer sequences, expected product length, and gene accession are given in Table 2.

Western Blot

3 SF- or TGFβ2-treated samples were pelleted and pooled together from 3 different pTM samples within the same condition. To separate membrane proteins from heavier cellular debris the pooled cell pellets were homogenized in a hypotonic lysis buffer (20mM TRIS-HCl, 3mM MgCl2, 10mM NaCl, 10mM PMSF, 0.5 mM DTT, 20 mM NaF, 2 mM NaV, 0.5 µg/mL leupeptin) before centrifuging at 300x g for 5 minutes (4 °C). The resulting supernatant was removed and centrifuged again at >12,500 rpm for 30 minutes to pellet membrane proteins, which were then resuspended in RIPA Buffer (Santa Cruz). Proteins were separated on a 10% SDS-PAGE gel and transferred to polyvinylidene difluoride membranes (Bio- Rad). Membranes were blocked with 5% skim milk/2% BSA in TBST and incubated at 4 °C overnight with a primary antibody against TRPV4 (1:250, Alomone Labs #ACC-034) or rabbit antibody against β-tubulin (1:2000, Abcam #EPR1330). Appropriate secondary antibodies conjugated to HRP were used to visualize protein expression on an iBright CL750 imaging system (Thermo Fisher Scientific). β-Tubulin expression was used to standardize protein levels between samples.

Calcium Imaging

pTM cells were seeded onto Collagen-I (Corning) coated coverslips and cultured in TMCM media (ScienCell) as described (39, 41). The cells were serum starved for 24 hours followed by serum-free TMCM with or without TGFβ2 (1 or 5 ng/mL) for 24 hours or five days. The cells were loaded with 10 µM of the ratiometric indicator Fura-2 AM (Kd at RT = 225 nM (Invitrogen/ThermoFisher) for 30-60 minutes. Coverslips were placed in a RC-26G chamber platform (Warner Instrument Corp) and perfused with external saline (pH 7.4) (in mM): 80 NaCl, 4.7 KCl, 1.2 MgCl2, 10 D-Glucose, 19.1 HEPES sodium salt, 2 CaCl2 and osmolality adjusted to 300 mOsm using D-mannitol. External solutions were delivered via a manually controlled gravity-fed eight-line manifold system, with perfusion speed kept constant to minimize changes in shear. Epifluorescence imaging was performed using an inverted Nikon Ti microscope with a 40x 1.3 N.A. oil objective and Nikon Elements AR software. 340 nm and 380 nm excitation were delivered by a high-intensity 150W Xenon arc lamp (Lambda DG-4; Sutter Instruments), high pass-filtered at 510 nm and detected with a 12-bit Delta Evolve camera (Photometrics/Teledyne).

GSK101 (10 nM) evoked Δ[Ca2+]i was assessed as ΔR/R (dividing the difference between the peak GSK- evoked F340/F380 signal during stimulation and baseline F340/F380 signal by the baseline F340/F380 signal).

Every data point represents a separate experimental day and pTM cell strain, each with 3-5 control and 3- 5 TGFβ2-treated slides tested on the same day. TGFβ2 datapoints represent the average GSK101 evoked ΔR/R across all TGFβ2 cells as a % of the average ΔR/R of control cells from the same cell strain on the same day.

Collagen hydrogel contraction assay

Rat tail collagen type I (Corning, Thermo Fisher Scientific) was prepared at a concentration of 1.5 mg/ml according to the manufacturer’s instructions. Five hundred microliters of the hydrogel solution were pipetted into 24-well culture plates. Upon complete collagen polymerization, pTM cells were seeded at 1.5 × 105 cells/well atop the hydrogels and cultured in DMEM + 10% FBS + 1% PSG for 48 hours to facilitate even cell spreading. Next, constructs were cultured in serum-free DMEM + 1% PSG supplemented with: i) control (vehicle: 0.008 mM HCl + 0.0004% BSA; 0.025% DMSO), ii) TGFβ2 (5 ng/ml; R&D Systems), or iii and iv) TGFβ2 + HC067047 (5 µM in DMSO) for 36 hours before carefully releasing the hydrogels from the walls using a sterile 10 µl pipette tip to facilitate contraction. The next morning, fresh serum-free DMEM + 1% PSG supplemented with 0.0025% DMSO (= vehicle) was added to groups i-iii; group iv received serum-free DMEM + 1% PSG supplemented with GSK1016790A (25 nM in DMSO). Plates were longitudinally imaged at 600 dpi resolution with a CanoScan LiDE 300 flatbed scanner (Canon USA) at 0, 15, 30, 60, and 120 minutes. Hydrogel construct size was quantified using FIJI software (National Institutes of Health) (127).

Electrophysiology

Borosilicate patch-clamp pipettes (WPI) were pulled using a P-2000 horizontal micropipette puller (Sutter Instruments), with a resistance of 6-8 MΩ. The internal solution contained (mM): 125 K-gluconate, 10 KCl, 1.5 MgCl2, 10 HEPES, 10 EGTA, pH 7.4. Patch clamp data was acquired with a Multiclamp 700B amplifier, pClamp 10.6 software and Digidata 1440A interface (Molecular Devices), sampled at 5kHz and analyzed with Clampfit 10.7. Current-voltage relationships were assessed using Vm steps from -100 to + 100 mV against a holding potential of -30 mV. Current density was measured as the average current during GSK101 exposure subtracted by the average current from the same cell during baseline perfusion.

IOP Measurements

A TonoLab rebound tonometer (Colonial Medical Supply) was used to measure IOP of awake mice between 12-2 P.M. IOP was determined from the mean of 10-20 tonometer readings. Nocturnal measurements were conducted between 9-10 P.M. under 2.5% isoflurane delivered by a Somnosuite isoflurane vaporizer (Kent Scientific). After animals recovered from intracameral HC-06/PBS injections, IOP was measured daily. IOP was measured every day for 4-5 consecutive days to confirm a stable return to baseline. IOP data for individual cohorts was binned into weeks of experimental time to group values for analysis.

Lentiviral Transduction

Lentiviral stock for TGFβ2 (C226,228S) was purchased from VectorBuilder Inc. (VB170816-1094fnw, pLV[Exp]-CMV> {hTGFB2[NM_003238.3](C226,228S)}) (23). Scrambled control lentivirus was purchased from SignaGen Laboratories (LM-CMV-Null-Puro). Mice were anesthetized with an intraperitoneal IP injection of ketamine/xylazine (90 mg/10 mg/ kg body weight), followed by eyedrops containing 0.5% proparacaine hydrochloride and 1% tropicamide ophthalmic solution to numb the eyes and dilate the pupils. Anesthetized mice were secured to allow stereotaxic injection of lentivirus.

Intravitreal injections were conducted by creating a guide hole with a 30-gauge needle 1-2 mm equatorial of the cornea-scleral border followed by insertion of a 12° beveled 33-gauge Hamilton syringe (Hamilton Company) secured to a stereotaxic rig (World Precision Instruments) used to insert the needle 2-3 mm into the eye. Each eye was injected with a 2uL bolus of lentivirus diluted to 1x106 TU/µL over the course of one minute, before the needle was quickly drawn and the pilot hole treated with erythromycin ophthalmic ointment USP (Bausch & Lomb). The efficiency of LV-TGFβ2 OHT induction in WT animals was close to 100%. No differences in observable health post-injection were detected between wild type and Trpv4-/- animals or LV-Ctrl and LV-TGFβ2 injected animals.

Intracameral Microinjections

Mice were anesthetized and treated with eyedrops as above, before being placed on an isothermal heating pad. HC-06 (100 µM) or PBS with DMSO (0.5%) as a vehicle were injected into the anterior chamber using a blunt tip Hamilton syringe (Hamilton Company) through a guide hole made using a 30-gauge needle. At the end of each injection a small air bubble was introduced to seal the cornea and minimize fluid outflow. 0.5% Erythromycin ophthalmic ointment USP (Bausch & Lomb) was applied to the eye after the procedure. Intracameral injections were not associated with observable inflammation, corneal opacity or behavioral changes. For the nocturnal IOP experiments in Figure 6, two animals were injected with PBS while two were injected with HC-06. When OHT was stably reestablished a week post- injection, the treatment groups were switched, and experiments repeated to obtain four eyes/treatment group for Figure 6C-D.

Statistical Analysis

GraphPad Prism 9 was used for statistical analysis. Means are plotted ± SEM unless otherwise noted. One-sample t-tests were used to determine whether TGFβ2 treated groups were significantly different than untreated control groups, while one-way ANOVA or two-way ANOVA along with Tukey or Bonferroni’s multiple comparisons test were used to compare multiple groups.

eLife Assessment

This fundamental work extends our understanding of the role of TGFβ2 as a modulator of mechanosensing in the eye and identifies the TRPV4 ion channel as a common regulator of Trabecular Meshwork (TM) contractility and pathological OHT. The data and evidence are convincing, with some minor limitations. This work will clearly be of interest to researchers investigating the role of mechanosensors in the TM and may underpin future research into treatments that aim to lower intra ocular pressure. This work will additionally be of interest to the growing field of researchers investigating the regulation of force sensing via ion channels and their roles in health and disease, in particular the ion channel TRPV4.

Reviewer #1 (Public review):

Summary:

This comprehensive study employed molecular, optical, electrophysiological and tonometric strategies to establish the role of TGFβ2 in transcription and functional expression of mechanosensitive channel isoforms alongside studies of TM contractility in biomimetic hydrogels, and intraocular pressure regulation in a mouse model of TGFβ2 -induced ocular hypertension. TGFβ2 upregulated expression of TRPV4 and PIEZO1 transcripts and time-dependently augmented functional TRPV4 activation. TRPV4 activation induced TM contractility whereas pharmacological inhibition suppressed TGFβ2-induced hypercontractility and abrogated ocular hypertension in eyes overexpressing TGFβ2. Trpv4-/- mice resisted TGFβ2-driven increases in IOP. These data establish a fundamental role of TGFβ as a modulator of mechanosensing and identifies TRPV4 channel as a common mechanism for TM contractility and pathological ocular hypertension.

Strengths:

The manuscript is very well written and details the important function of TRPV4 in TM cell function. These data provide novel therapeutic targets and potential for disease-altering therapeutics.

Weaknesses:

The experimental rigor and design of the noctural IOP experiments was weak with low n values and differing methods of IOP measurement (conscious versus anesthetized). The same method of IOP measurement needs to be used for all measurements to make any conclusions on the circadian patterns of IOP in each condition.

Reviewer #2 (Public review):

The manuscript by Christopher N. Rudzitis et al. describes the role of TGFβ2 in the transcription and functional expression of mechanosensitive channel isoforms, alongside studies on TM contractility in biomimetic hydrogels and intraocular pressure. Overall, it is a very interesting study, nicely designed, and will contribute to the available literature on TRPV4 sensitivity to mechanical forces.

I have the following comment for the authors to address.

Figure 1A-C.

Often there is a difference between the massage and transcript data. I recommend the authors to confirm with qPCR data with another mode of protein measurements.

Does direct TRPV4 activation also induce the expression of these markers? Does inhibition of TRPV4, after TGF-β treatment, prevent the expression of these markers? Is TRPV4 acting downstream of this response?

Figure 1D. Beta tubulin is not a membrane marker. Having staining of b tubulin in membrane fraction shows contamination from the cytoplasm.

Does the overall expression also increase?

Figure 4A: it is not very clear. I recommend including a zoom image or better resolution image.

Figure 5B and 6B.

Why there is a difference between groups in pre-injection panel. As Figure 5A, in pre-injection, there is no difference between LV-TGFβ and LV-control while in 5B there is a significant difference between these groups.

Discussion section.

Line 279, . "TRPV4 channels in cells treated with TGFβ2 are likely to be constitutively active" ... needs to be discussed further.

Line 280: "The residual contractility in HC-06-treated cells may reflect TGFβ2-mediated contributions from Piezo1."

Piezo1 has a low threshold for mechanosensitivity. How do the authors discuss the observation that, in the presence of Piezo1, TRPV4 has a more prominent mechanosensory function? Is this tied to TGFβ signalling?

Author response:

The following is the authors’ response to the original reviews

Reviewer #1 (Public review):

The experimental rigor and design of the noctural IOP experiments was weak with low n values and differing methods of IOP measurement (conscious versus anesthetized). The same method of IOP measurement needs to be used for all measurements to make any conclusions on the circadian patterns of IOP in each condition.

One of the goals of our study was to confirm the results from the Patel et al (2021; PMID33853948) study, which in which nocturnal IOP measurements were conducted in anesthetized mice and diurnal IOP measurements in awake animals but we agree with both Reviewers that IOP should be measured under identical experimental conditions. Parenthetically, the number of animals per each treatment paradigm in the original version (N = 4) was sufficient to produce statistical significance for diurnal control vs diurnal TGFB, and diurnal control vs nocturnal control conditions.

To address the comment, we generated an additional cohort of TGFb2-expressing mice (N = 6) in which nocturnal and diurnal measurements were performed in awake animals. The results are shown in the revised Figure 6. Similar to the anesthetized cohort, the diurnal IOP in Lv-TGFB2 mice was statistically indistinguishable from the nocturnal value, indicating that TGFB2-induced OHT is not additive to physiological (circadian) OHT. The TRPV4-dependence of ocular hypertension induced by physiological and pathological methods suggests that the channel functions as a final common mechanism for ocular hypertension.

Reviewer #2 (Public review):

Figure 1A-C. Often there is a difference between the massage (message?, op. authors) and transcript data. I recommend the authors to confirm with qPCR data with another mode of protein measurements.

We are not sure we understand the Reviewer’s comment regarding the “difference between the message and transcript data” but note that the mRNA data shown in panels A & B are confirmatory of previously published transcriptomic and proteomic screens (eg, Fleenor et al., IOVS 2006; Bollinger et al., IOVS 2011; Callaghan et al., Scientific Reports 2022; Li et al., Current Eye Research 2022 etc) and were included to show that the transcriptional response of canonical SMAD and pro-fibrotic genes unfolds as predicted from previous work. With regard to TRPV4 signaling, we expand transcriptomic data with protein analysis (Western blots) and functional analyses (measurements of TRPV4-mediated current and calcium imaging). Transcriptomic, protein expression, electrophysiological and imaging experiments revealed a remarkable consistency in TGFB2-dependence of gene (Fig. 1C) and protein expression (Fig. 1D), transmembrane current (Fig. 3C) and intracellular calcium (Fig. 2).

Parenthetically, we attempted to get a sense for the TGFB2-dependence of Piezo1 protein expression by conducting Western blots with multiple antibodies and experimental conditions. These efforts were unsuccessful, presumably due to the complexity (30-40 TM domains) and large molecular weight (280-300 kDa) of the protein. We note, however, that Piezo1 signaling cannot account for the observed OHT given that studies by us and others (Yarishkin et al., 2021, PMID: 33226641 and Zhu et al., 2021; PMID: 33532718) associated Piezo1 signaling with facility increases. The revised m/s reads: “The suppression of outflow facility by Piezo1 inhibitors applied under in vitro and in vivo conditions (39, 81) instead suggests that Piezo1 opposes the hypertensive functions of TRPV4.” The preprint by Redmon et al. (bioRxiv 2024, PMID 39041037) expands the TRPV4-dependence of OHT to microbead-induced, steroid-induced and nocturnal models of OHT to indicate that TRPV4 functions as a universal driver of elevated IOP. We reiterate this in the revised Discussion.

Does direct TRPV4 activation also induce the expression of these markers? Does inhibition of TRPV4, after TGF-β treatment, prevent the expression of these markers? Is TRPV4 acting downstream of this response?

A RNASeq study conducted by us (Rudzitis et al., under review) suggests that the agonist GSK101 has minimal effect on the fibrotic and canonical pathways shown in panels A and B. These data are beyond the scope of the present study. They will be published elsewhere, however, we include the data associated with genes depicted in panels A and B for the reviewer at the end of this Response.

We conducted an additional series of experiments to test whether TGFB2-induced upregulation of the TRPV4 and Piezo1 genes is itself TRPV4-dependent. As shown in the new SFig. 1, upregulation of the two genes is unaffected by TRPV4 inhibition.

Figure 1D. Beta tubulin is not a membrane marker. Having staining of b tubulin in membrane fraction shows contamination from the cytoplasm. Does the overall expression also increase?

b-tubulin associates with the plasma membrane by binding to integral membrane proteins in the plasma and organellar membranes through palmitoylation and attachment to linker proteins and as an integral component of exocytotic vesicles (Wolff, BBA 2009; Hogerheide et al., PNAS 2017). The protein is often used as a loading control for the TRPV4 protein (please see https://www.cellsignal.com/products/primary-antibodies/trpv4-antibody/65893; Grove et al., Science Signaling 2019 and Moore et al., PNAS 2013). Parenthetically, our RNASeq studies did not find modulation of b-tubulin expression by TGFβ2 [CNR and DK, unpublished observations].

We examined the overall (cytosolic and membrane) TRPV4 expression and observed, similarly to the membrane fraction alone (Figure 2), upregulation following cytokine stimulation:

Western blot, total protection extract from control and TGFb2-treated TM cells [Alomone antibody].

These results in our estimation do not add to the overall narrative and were not included into the paper.

Figure 4A: it is not very clear. I recommend including a zoom image or better resolution image.

We include a whole-page image as the new SFigure 4.

Figure 5B and 6B. Why there is a difference between groups in pre-injection panel. As Figure 5A, in pre-injection, there is no difference between LV-TGFβ and LV-control while in 5B there is a significant difference between these groups.

We revised Figure Legends to clarify that “pre-injection” in Figures 5B and 6B refers to IOP measurements before the intracameral injection of HC-06 not pre-injection of lentiviral constructs.

Discussion section. Line 279: "TRPV4 channels in cells treated with TGFβ2 are likely to be constitutively active" ... needs to be discussed further.

We rewrote the paragraph to clarify that TRPV4 is a thermosensitive channel that is expected to be constitutively active at the incubator temperature:

“The effectiveness of TRPV4 inhibition in suppressing TGFB2-induced contractility (Fig. 4) is consistent with constitutive activation of TRPV4 channels in incubator-cultured cells. TRPV4 is a thermosensitive channel (Q10 ~10). Mouse TRPV4 is activated by physiological temperatures (Chung et al., 2003; Shibasaki et al., 2007) with peak activation between ~34 - 37oC (Guler et al., 2003). The several-fold increase in functional expression of the channel in TGFB2-treated cells (Fig. 2) would be expected to promote tonic influx of Ca2+ and Ca2+-dependent cellular signaling. The abrogation of the contractile response in the presence of HC-06 indicates that TRPV4-mediated Ca2+ influx represents the principal source of calcium that drives the contractile response. Consistent with this, supplementation with the agonist GSK101 was sufficient to evoke TM contraction (Fig. 4B).”

Line 280: "The residual contractility in HC-06-treated cells may reflect TGFβ2-mediated contributions from Piezo1." Piezo1 has a low threshold for mechanosensitivity. How do the authors discuss the observation that, in the presence of Piezo1, TRPV4 has a more prominent mechanosensory function? Is this tied to TGFβ signalling?

This is an interesting question. Our macroscopic and single channel recordings of Piezo1 activity in TM cells recapitulate the time course published in the original Coste et al. (2010) study, showing the channel inactivates within 10-100 msec (Yarishkin et al., 2021). Thus, it is likely that the channel is largely inactivated during chronic ocular hypertension. Indeed, it has been suggested that resting membrane tension alone may be sufficient to inactivate Piezo1 (Lewis and Grandl, 2015), with cells grown on stiff substrates (e.g., under our experimental conditions) experiencing almost complete Piezo1 inactivation. We propose that the primary function of Piezo channels may be to sense and transduce transient mechanical loading. The remarkable IOP-lowering effectiveness of TRPV4 antagonists and knockdown indicates that - in contrast to Piezo1 - TRPV4 activation is sustained.

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

The complete strain name for the Trpv4-/- mice are missing.

Corrected.

The layout for Figure 6 is confusing as HC-06 was only used in panels B and C but the labels are above panel A.

Corrected.

Reviewer #2 (Recommendations for the authors):

Only two mice were used for the noctural IOP experiments. Justification for retreating the same mice in opposite eyes and counting it as n=4 is not rigorous or justified.

The number of mice investigated in the original submission was four. In Week 1, two mice underwent PBS injections and 2 two mice were treated with HC-06. After the baseline was re-established in Week 2, the treatments were reversed.

We supplemented these numbers with an additional cohort of 6 mice, with identical results re: nocturnal vs diurnal IOP. These data are presented in the revised Figure 6.

Why are daytime IOPs measured in awake mice but noctural IOP's measured in isoflurane anesthetized mice? Anesthesia is well known to effect IOP and using two different methods could alter the results, especially when comparing between the groups. This could be why you did not see a noctural rise in the TGFB injected eyes. The same method needs to be used for all measurements to make any conclusions on the circadian patterns of IOP in each condition.

This is a good point, please see our response above.

Version 2

Abstract

Abstract

The risk for developing primary open-angle glaucoma (POAG) correlates with the magnitude of ocular hypertension (OHT) and the concentration of transforming growth factor-β2 (TGFβ2) in the aqueous humor. Effective treatment of POAG requires detailed understanding of interaction between pressure sensing mechanisms in the trabecular meshwork (TM) and biochemical risk factors. Here, we employed molecular, optical, electrophysiological and tonometric strategies to establish the role of TGFβ2 in transcription and functional expression of mechanosensitive channel isoforms alongside studies of TM contractility in biomimetic hydrogels, and intraocular pressure (IOP) regulation in a mouse model of TGFβ2 -induced OHT. TGFβ2 upregulated expression of TRPV4 and PIEZO1 transcripts and time-dependently augmented functional TRPV4 activation. TRPV4 agonists induced contractility of TM-seeded hydrogels whereas pharmacological inhibition suppressed TGFβ2-induced hypercontractility and abrogated OHT in eyes overexpressing TGFβ2. Trpv4-deficient mice resisted TGFβ2-driven increases in IOP, but nocturnal OHT was not additive to TGFβ-evoked OHT. Our study establishes the fundamental role of TGFβ as a modulator of mechanosensing in nonexcitable cells, identifies the TRPV4 channel as the final common mechanism for TM contractility and circadian and pathological OHT, and offers insights for future treatments that can lower IOP in the sizeable cohort of hypertensive glaucoma patients that resist current treatments.

Introduction

Primary open-angle glaucoma (POAG), an irreversible blinding disease, afflicts ∼3.5% of the global population (1). Its incidence and severity are proportional to the amplitude and duration of ocular hypertension (OHT) (2, 3), which correlates with retinal ganglion cell dysfunction, neuroinflammation, and oxidative stress (4, 5). Biomechanical factors, glucocorticoids, and the cytokine transforming growth factor-β2 (TGFβ2) contribute to POAG by compromising the funneling of aqueous humor (AH) from the trabecular meshwork (TM) into Schlemm’s canal (SC). Elevated IOP enhances the contractility of the juxtacanalicular TM (JCT), a circumocular tissue comprised of extracellular matrix (ECM) beams populated by mechanosensitive and smooth muscle-like cells- thereby increasing AH outflow resistance. The molecular mechanism that links TM pressure sensing to the contractile response is not known but is likely to underpin the tissue’s sensitivity to compressive, tensile, osmotic, shear and traction forces which collectively regulate expression of numerous TM genes and secretion of ECM proteins (6–12).

The increase in trabecular outflow resistance induced by mechanical stress, glucocorticoids, and TGFβ2 manifests through two distinct components: a dynamic, reversible phase amenable to cytoskeletal and Rho kinase inhibition, and a chronic phase, characterized by transdifferentiation of TM cells into fibrotic and contractile myofibroblasts (16–18). TGFβ2-induced fibrotic remodeling has been linked to POAG: (i) TM cells derived from POAG patients secrete more active TGFβ2 compared to cells isolated from healthy donors (19), (ii) the risk of POAG is proportional to [TGFβ2]AH (20–22), and (iii) ectopic ocular expression of TGFβ2 suffices to induce OHT (23, 24), likely via aberrant secretion of ECM proteins and enhanced TM contractility (25, 26). The cognate TGFβ1 isoform induces similar fibrotic responses in fibroblasts, epithelial, and endothelial cells from heart, kidney, skin, and/or lung, suggesting induction of conserved fibrogrenic programs (27–30). However, the contribution to OHT by ocular TGFβ expression cannot be disambiguated from the changing biomechanical environment: TGFβ release is activated by tissue contractility and tension (31, 32), and TGFβ activity correlates with mechanical stress gradients which may drive a cellular epithelial-mesenchymal transition-like phenotype (EMT; 33-35).

Despite its clinical relevance, our understanding of TM mechanotransduction and its contribution to IOP homeostasis remains rudimentary. Strain and shear stress have been hypothesized to engage primary cilia and integrins, as well as mechanosensitive TRPV4, Piezo1 and TREK-1 channels (36–39), yet it remains unclear whether these mechanosensors regulate TM contractility, are influenced by POAG inducers like TGFβ2 or glucocorticoids, or contribute to chronic fibrosis. Among these, TRPV4 (Transient Receptor Potential Vanilloid isoform 4), a tetrameric channel with PCa/PNa ∼ 10 (40), is strongly expressed in rodent and human TM (36, 51) where it carries the principal component of the pressure-activated transmembrane current and is activated by stretch, shear and swelling (7, 10, 37, 39, 41, 42). Pharmacological inhibition and deletion of the TRPV4 gene alter pressure gradients in the brain, kidney, lung, and bladder (46–50). TRPV4 mutations underpin sensorimotor neuropathies, skeletal dysplasias, retinal degeneration and ocular dysfunction (43–45) while the role of TRPV4 signaling in OHT remains contentious, with evidence suggesting IOP-lowering as well as IOP-elevating effects. TRPV4-dependence of conventional outflow has been linked to diverse downstream effector mechanisms (e.g., eNOS and RhoA activation, phospholipid-cholesterol-caveolin regulation, OCRL inositol-5-phosphatase interaction, modulation of cell-ECM contacts, polyunsaturated fatty acid release, and Piezo1 signaling; 7, 37, 41, 52–55) and thus leads towards testable hypotheses: if TM-intrinsic TRPV4 sustains steady-state normotension, promotes outflow via eNOS-dependent TM relaxation and mitigates TGFβ2 -driven fibrosis (7, 52), TRPV4 inhibition should induce OHT. Conversely, if TRPV4 exacerbates OHT, its blockade and deletion should reduce IOP.

In this study, we tested these hypotheses through investigation of reciprocal TRPV4- TGFβ2 interactions that perpetuate the vicious feedback loop between mechanical stressors, TM contractility, and OHT. We demonstrate that inhibition and deletion of TRPV4 lower IOP in TGFβ2 overexpression-induced and circadian OHT models and suppress TM contractility in TGFβ2-treated biomimetic hydrogels. The cytokine promoted upregulation of EMT-associated genes alongside increased transcription and activity of TRPV4, potentially sensitizing TM cells to physiological mechanical cues. Although TRPV4 activity was required to maintain OHT under physiological (nocturnal) and pathological (cytokine-induced) conditions, their respective IOP elevations were not additive, suggesting convergence on a shared common pathway. Collectively, these findings position TRPV4 as a critical nexus of TGFβ2 -induced TM contractility and IOP dysregulation. As such, TRPV4 perpetuates the vicious feedback loop between mechanical stressors and TM contractility and thus represents an ideal therapeutic target in glaucoma cases that resist current treatments.

Results

TGFβ2 drives overexpression of genes that encode fibrotic markers and mechanosensitive ion channels

Human TM cells respond to TGFβ2 with increased biosynthesis, deposition and degradation of ECM, altered autophagy, upregulation of F-actin stress fibers, a-smooth muscle actin (aSMA) (19, 25, 26, 56, 57), but it is unclear whether cells undergoing TGFβ2-induced fibrotic remodeling also exhibit altered capacity for sensing and transduction of mechanical stimuli. We profiled genes that encode known TM mechanochannels together with a selection of key cytoskeletal, ECM, and fibrotic markers in primary TM cells (pTM) isolated from 3-7 donors without history of visual dysfunction (Figure 1 A - C ). Five-day exposure of pTM cells to a physiological concentration of TGFβ2 (1 ng/mL) increased the expression of EMT-promoting transcription factor SNAI1 (SNAIL1, P = 0.0094) and fibronectin (FN1, P = 0.0263), while expression of connective tissue growth factor 2 (CCN2, alternatively CTGF) was elevated in 5/5 pTM cell strains without reaching significance (P = 0.0909). Expression of fibroblast-specific protein 1 (FSP1, a calcium-binding fibroblast marker), yes-associated protein 1 (YAP1, a stiffness induced hippo-pathway transcription factor) and ACTA2 (αSMA, associated with cell contractility) was not consistently impacted by TGFβ2 while transcription of myocilin (MYOC) decreased across 4/4 pTM strains (P = 0.0055) (Figure 1 B ). Indicative of feedback inhibition (58), TGFβ2-treatment downregulated transcript levels of transforming growth factor beta receptor 2 (TGFBR2, P = 0.0219) and upregulated expression of autoinhibitory SMAD family protein 7 (SMAD7, P = 0.0461) without affecting SMAD2 or SMAD3 expression. TGFβ2 thus promotes selective upregulation of ECM and fibrosis-related genes together with cell dedifferentiation and activation of autoregulatory SMAD mechanisms.

TGFβ2 induces a fibrotic phenotype in pTM cells and increases expression and membrane insertion of the TRPV4 channel.

(A-B) Five-day TGFβ2 treatment (1ng/mL) significantly altered expression of TGFβ pathway effectors, cytoskeletal machinery, and canonical fibrotic markers. (C) TGFβ2 treatment significantly increased TRPV4 and PIEZO1 expression, but not TREK1 and TRPC1 expression. Mean ± SEM shown. N = 4 - 8 experiments, each gene tested in 3-7 different pTM strains (See Table 1). Two-tailed one sample t-test of TGFβ2-induced gene expression levels as a percent of control samples. (D) Isolation of membrane proteins from two separate pooled pTM samples suggests TGFβ2 treatment drives increased TRPV4 membrane insertion. N = 2 independent pooled samples, 3 pTM strains were pooled per sample. * P < 0.05, ** P < 0.01.

Analysis of genes encoding mechanosensitive channels implicated in outflow modulation (36, 39, 59, 60) showed a 102.5% increase in expression of TRPV4 (P = 0.0193) and 78.9% increase in PIEZO1 expression (P = 0.0114) across 8 replicates including 7 distinct pTM strains. (Figure 1 C ). Conversely, TGFβ2 exposure did not affect expression of the TRPC1 gene (P = 0.261) and had variable, strain-dependent effects on transcript levels of KCNK2 (P = 0.293, encoding the TREK-1 channel). Thus, TGFβ2 promotes selective transcriptional upregulation of genes that encode a subset of mechanosensitive proteins alongside fibrotic upregulation and cell dedifferentiation. Finally, we tested whether TGFβ2-induced upregulation of TRPV4 and Piezo1 is TRPV4-dependent, however, inclusion of the selective TRPV4 inhibitor HC067-47 (HC-06; 5 mM) had no effect on transcriptional upregulation compared to TGFβ2 treatment alone (SI Appendix Figure S1).

TGFβ2 exposure time-dependently augments TRPV4-mediated current and Δ[Ca2+]i

To assess the functional relevance of TGFβ2-dependent transcriptional upregulation we determined the membrane expression and functional activation of TRPV4, which mediates the pressure-activated current and calcium signaling, regulates cytoskeletal dynamics and modulates conventional outflow resistance in vitro (37, 41). TGFβ2 exposure produces an increase in levels of membrane-bound TRPV4 protein (Figure 1 D ) in western blot of two grouped pTM membrane protein samples. While low amounts of TRPV4 were visible in the membrane fractions in control samples, TGFβ2 treatment produced an increase in the higher weight TRPV4 band, suggesting there may be isoform-specific TGFβ2-induced responses and increased TRPV4 translation leading to elevated TRPV4 trafficking, membrane insertion and/or lipid raft interaction (52).

Functional expression was assessed by tracking [Ca2+]i changes in cells exposed to the selective TRPV4 agonist GSK1016790A (GSK101, 10 nM) using ratiometric Fura2-AM Ca2+ dye, with TGFβ2-treated and control cells tested on the same day. All pTM strains responded to GSK101 with robust [Ca2+]i increases which reached peak within 5 min before the majority of responding cells gradually decreased to a steady plateau (Figure 2 C ). TGFβ2-treated cells exhibited a remarkable potentiation of GSK101-evoked [Ca2+]i responses compared to control cells, with 5/5 cell strains showing an increase in the Δpeak/baseline F340/F380 response equivalent to 258.4% ± 61.7% of the control response in (P = 0.0046) (Figure 2 A-B ). The fraction of GSK101 responders and the overall time course of responses between groups were not significantly different, indicating that TRPV4 potentiation primarily affects TRPV4-expressing cells. Thus, TGFβ2 treatment promotes TRPV4 expression and functional activity.

TRPV4-mediated Ca2+ influx is potentiated by five-day TGFβ2 treatment.

(A) Five-day TGFβ2 treatment (1 ng/mL) increased TRPV4 agonist-induced (GSK101, 10 nM) Ca2+ influx in pTM cells compared to serum-free media alone treated cells tested on the same day (N = 5 pTM strains, n = 3 - 5 slides/condition/day, individual data points over mean ± SEM). Two-tailed one sample t-test of TGFβ2-treated cell average GSK101 response as a percent of control samples from the same pTM strain on the same day. (B) Violin plots showing the distribution of GSK101-induced Ca2+ responses for each pTM strain tested in A. Thick dashed line indicates mean, while light dashed line indicates quartiles. (C) Representative traces showing TRPV4 agonist-induced Ca2+ influx (seen as an increase in F340/F380) in pTM (mean ± SEM of 4 representative cells/ group), alongside example Fura-2-loaded pTM cells before (i), during (ii), and after (iii) GSK101 application. Scale bar = 50 µm. ** P < 0.01

To gain insight into the time- and dose-dependence of TGFβ2-dependent TRPV4 signaling modulation pTM cells were treated for 24 hours, at 1 ng/mL and 5 ng/mL concentrations of TGFβ2. GSK101-stimulated Ca2+ influx was not significantly increased by 24-hour TGFβ2 treatment at 1 ng/mL (Δpeak/baseline F340/F380 = 117.0% ± 23.6% of control) or 5 ng/mL (Δpeak/baseline F340/F380 = 133.6% ± 34.5% of control) (Figure 3; SI Appendix Figure S2); the potentiation of both was significantly lower relative to the five-day 1 ng/mL TGFβ2 treatment (P < 0.0011; Figure 3 A ). GSK101 evoked a moderately outwardly rectifying nonselective current (IGSK-Ibaseline) with reversal potential at ∼0 mV (Figure 3 C ). While its amplitude was variable, mean current density consistently increased in cells treated for 1 day with TGFβ2 (n = 10; 5 ng/mL) relative to the control group (n = 11). The potentiating effect of TGFβ2 on TRPV4 activity thus appears to be time-dependent but is significant after chronic exposure to relatively low-dose TGFβ2.

TGFΒ2-induced TRPV4 potentiation is not seen at a shorter time period, regardless of treatment strength.

(A) TGFβ2 treatments for 24 hours at 1 ng/mL (N = 6 pTM strains, n = 3 - 5 slides/condition/day) or 5 ng/mL (N = 5 pTM strains, n = 3 - 5 slides/condition/day) did not show potentiation of GSK101-evoked TRPV4 Ca2+ influx (SI Appendix Figure S2) and were significantly lower than cells treated with TGFβ2 for 5d at 1ng/mL (5d TGFβ2 results from Figure 2A). Individual data points over mean ± SEM. One-way ANOVA with Tukey’s multiple comparisons test, statistics for individual 1d treatment groups compared to control groups shown in Figure S1. (B) Representative traces for GSK101 response following 24-hour TGFβ2 treatment, traces show mean ± SEM of 3-4 cells. (C) Average current density in response to GSK101 (24-hour control: n = 11 cells, 24-hour TGFβ2: n=10 cells) shows generally increased current in TGFβ2-treated cells. Data shows mean ± SEM (D - E) Violin plots of individual cell strains shown in A. Thick dashed line indicates mean, while light dashed line indicates quartiles. ** P < 0.01, *** P < 0.001

TGFβ2-Induced TM contractility requires TRPV4 activation

The IOP-lowering effectiveness of Rho kinase inhibitors and latrunculins (57, 61–63) indicates that sustained increases in outflow resistance require tonic actin polymerization and contractility. TGFβ2 drives the TM myofibroblast contractile response (57) while the role of mechanosensation remains unknown. To ascertain whether TRPV4 upregulation (Figures 1-2) contributes to the contractile response, we seeded pTM cells into high-compliance Type I collagen hydrogels (57) (Figure 4, SI Appendix Figure S3-4). Hydrogels that were incubated with TGFβ2 (5 ng/mL) showed profound increases (P < 0.0003) in the rate and the magnitude of contraction at all time points (Figure 4, SI Appendix Figure S3-4). Simultaneous treatment with HC-06 (5 µM) significantly reduced the extent of TGFβ2-induced TM contractility (P < 0.0001). To determine whether TRPV4 activation is sufficient to induce the contractile response, the antagonist was washed out and hydrogels supplemented with GSK101 (25 nM). 15 minutes post-treatment, the constructs responded to the agonist with transient contraction (Figure 4 C ; SI Appendix Figure S3, P < 0.01), with a time course mirroring GSK101-induced [Ca2+]i elevations (Figures 2-3). The effects of TRPV4 inhibition and activation were consistent across all pTM strains tested (N = 3). TRPV4-mediated Ca22+ influx is therefore sufficient to induce TM contractility and necessary for pTM hypercontractility induced by TGFβ2.

Effects of TRPV4 inhibition/activation on TGFβ2-induced contraction of TM cells.

(A) Representative longitudinal 24-well plate scans of collagen type I hydrogels seeded with pTM subjected to the different treatments (dashed lines outline size of contracted constructs). (B) Longitudinal quantification of hydrogel construct size compared to the control group at the 0 minute time point. (C) Detailed comparisons between groups at each experimental time point. n = 6 hydrogels/group. One-way ANOVA with Tukey’s multiple comparisons test, data in (B & D) shows individual data points over mean ± S.D. One pTM strain shown: TGFβ2-induced contractility induction, HC-06-mediated rescue of hypercontractility, and GSK101-induced transient (15 min) contraction were consistent across (3/3) pTM strains tested (SI Appendix Figure S3). ** P < 0.01, *** P < 0.001, **** P < 0.0001.

TRPV4 activation is required to maintain TGFβ2-induced OHT

To test whether TRPV4 contributes to TGFβ2 induced ocular hypertension (OHT) in vivo, we utilized the lentiviral TGFβ2 overexpression model developed by Patil et al. (23). Adult C57BL/6J mice (N = 5) were intravitreally injected with lentivirus overexpressing constitutively active human TGFβ2 (LV-TGFβ2). LV-TGFβ2-injected eyes, but not the contralateral eyes injected with a lentivirus containing a scrambled transgene (LV-Ctrl), exhibited significant IOP elevations one-week post-transduction (Figure 5 A , Week 2, ΔTGF-Ctrl = 4.0 mm Hg, P = 0.0143). By 2 weeks post-transfection, IOP in LV-TGFβ2 eyes reached 19.9 ± 4.7 mm Hg whereas IOP in LV-Control eyes remained at control levels (14.0 ± 1.2 mm Hg), with ΔTGF-Ctrl = 5.9 mm Hg (P = 0.0002). IOP remained elevated throughout the 4 weeks after the injection (Week 5, ΔTGF-Ctrl = 4.9 mm Hg, P = 0.0008). HC-06 (100 µM) microinjection into the anterior chamber of LV-TGFβ2 and LV-Ctrl eyes lowered IOP in LV-TGFβ2 eyes to 12.2 ± 1.7 mm Hg after 24 hours (ΔpostHC-preHC = -5.8 mm Hg) with no difference observed in IOP from LV-Ctrl eyes (12.6 ± 1.9 mm Hg, ΔpostHC-preHC = -0.3 mm Hg). LV-Ctrl eyes remained close to pre-injection levels post-HC-06 treatment (Figure 5 A - B ). IOP in LV-TGFβ2 eyes returned to hypertensive levels by 1-week post-HC-06 injection (Week 6-7, ΔTGF-Ctrl = 3.9 mm Hg, P = 0.0201). To determine the effect of the bolus injection alone, LV-TGFβ2 and LV-Ctrl eyes were reinjected with PBS 2 weeks after re-establishing the OHT baseline. The sham injection transiently reduced IOP in LV-TGFβ2 (ΔpostPBS-prePBS= -4.5 mm Hg) and LV-Ctrl (ΔpostPBSpre-PBS= -1.2 mm Hg) eyes; however, LV-TGFβ2 eyes returned to hypertensive levels by 48 hours post-injection (ΔTGF-Ctrl =3.6 mm Hg, p=0.0465) and to pre-injection levels after 72 hours (ΔTGF-Ctrl =5.4 mm Hg, p=0.0002). Bolus injection was less effective than HC-06 at all time points 24 hours post-injection (Week 8-9, Figure 5 B ). These data indicate that selective pharmacological inhibition of TRPV4 effectively and reversibly blocks TGFβ2-induced OHT.

TRPV4 activation is necessary to maintain LV-TGFβ2-induced ocular hypertension.

(A) Intravitreal injection of LV-TGFβ2 (week 1), but not LV-Control, elevates IOP in WT mice (N = 5 eyes/group) as early as one-week post-injection. Injection of TRPV4 antagonist HC-06, but not PBS, produced multiday IOP reduction in LV-TGFβ2 treated eyes. HC-06 and PBS injections did not affect IOP in LV-Control injected eyes. Two-way ANOVA with Bonferroni post-hoc analysis (B) Direct comparison of the results of PBS and HC-06 injections in the eyes shown in A. Two-way ANOVA with Bonferroni post-hoc analysis (C) Intravitreal injection of LV-TGFβ2 in Trpv4-/- mice (N = 6 eyes/group) resulted in only mild OHT; plotted against WT eyes at matching timepoints (3 WT cohorts including the 5 WT eyes shown in A-B, N = 8-15 eyes/group). (D) Statistical comparison of the IOP values shown in C. The IOP in LV-TGFβ2 WT eyes was significantly elevated compared to the LV-TGFβ2 Trpv4-/- eyes from 2 weeks post-injection. LV-Control injected eyes in WT or Trpv4-/- eyes remain close to the baseline value and are not significantly different. Two-way ANOVA with Bonferroni post-hoc analysis. (A, C) shows mean ± SEM. Data in (B, D) shows individual data points over mean ± SEM, * P < 0.05, ** P < 0.01, *** P < 0.001, **** P < 0.0001

To further evaluate the TRPV4-dependence of TGFβ-induced OHT we took advantage of mice with global Trpv4 knockdown (64–66). Trpv4-/- mice (N = 6) were injected with LV-TGFβ2 and LV-Ctrl vectors in contralateral eyes (Figure 5 C ). Additionally, two littermate control mice injected alongside the Trpv4-/-animals were added to previously collected WT LV-injected cohorts measured at the same timepoints (N = 8-15, Figure 5 C ). Pre-LV injection, IOP levels in Trpv4-/- animals were comparable to the WT cohort, indicating that TRPV4 activity does not regulate normotension. Similarly, IOP in LV-Ctrl-injected eyes was not significantly different between WT and Trpv4-/- animals at any point in the experiment (peak ΔCtrlKO-CtrlWT = -1.2 mm Hg, Figure 5 D , SI Appendix Figure S5). By two weeks post-injection (Week 3), LV-TGFβ2-treated Trpv4-/- eyes exhibited significantly lower IOP compared to the LV-TGFβ2 WT cohort (ΔTGFKO-TGFWT = -3.1 mm Hg, P = 0.0009, Figure 5 C ). While LV-TGFβ2 injected Trpv4-/- eyes exhibited mild OHT, the effect was significantly reduced compared to WT eyes and IOP decreased by two weeks post-injection (Figure 5 C - D ).

TGFβ2-induced and nocturnal OHT are non-additive but require TRPV4

Mammalian IOP is modulated by the circadian rhythm, with levels elevated at night and nocturnal IOP fluctuations implicated in POAG (7, 55, 67). We measured nocturnal (9:00-10:00 PM) IOP in LV-TGFβ2 (N = 4) and LV-Ctrl WT eyes (N = 4) from isoflurane anesthetized mice ∼ 2 months post-LV injection to determine whether nocturnal OHT is additive to TGFβ2-induced elevation observed during the daytime (12:00-2:00 PM, Figure 6 A ). LV-TGFβ2 injected eyes showed significant IOP elevation compared to LV-Ctrl eyes during daytime measurements (diurnal ΔTGFβ-Ctrl = 7.9 mm Hg, P < 0.0001) but the difference vanished at night (nocturnal ΔTGFβ-Ctrl = 0.2 mm Hg), indicating that TGFβ2-induced OHT is not additive to the circadian OHT. To test whether the IOP measurement was influenced by isoflurane-induced anesthesia, we repeated the nocturnal measurements in awake animals (N = 4-6 eyes/group). We observed no difference in nocturnal IOP between the two groups of animals was observed (Figure 6A-D , SI Appendix Figure S6). To determine whether physiological (nocturnal) OHT requires TRPV4 we microinjected LV-Ctrl (N = 4) and hypertensive LV-TGFβ2 (N = 4) eyes with PBS or HC-06. PBS injection did not affect IOP in LV-Ctrl or LV-TGFβ2 eyes at day or night (Figure 6 E-F ) except for a single LV-TGFβ2 eye that exhibited abnormally high nocturnal IOP (37 mm Hg) at the four-day timepoint. Conversely, HC-06 injection blocked LV-TGFβ2-induced IOP during the day (P < 0.001) and significantly lowered IOP in LV-Ctrl and LV-TGFβ2 eyes at night (∼5 mm Hg; P < 0.01). These data indicate that i) TRPV4 activation is necessary for OHT in the TGFβ2 overexpression mouse model (Figures 5-6) and the circadian IOP elevations ii) TGFβ2 -evoked OHT does not affect nocturnal IOP elevation in mice, and iii) TRPV4 inhibition does not disrupt the mechanisms that maintain daytime normotensive IOP (Figures 5-6).

TRPV4 inhibition inhibits nocturnal IOP elevation in control and TGFβ2 overexpressing eyes.

(A-B) Post-LV injection daytime (12-2:00 P.M) and nocturnal (9-10:00 P.M.) IOP compared in WT mice (N = 4-6 eyes/group) before drug treatment. LV-TGFβ2 eyes were elevated at daytime, but nocturnal OHT was not significantly different between LV-Ctrl and LV-TGFβ2 eyes in two separate cohorts of mice measured under isoflurane anesthesia (A) or while awake (B). (C-D) Anesthesia had no significant effect on measured nocturnal IOP. One-way ANOVA with Tukey’s multiple comparisons test. (E-F) PBS-injected eyes did not exhibit changes in daytime or nighttime intraocular pressure; however, HC-06 injection reduced TGFβ2-induced IOP elevations during the day and LV-Ctrl and LV-TGFβ2 nocturnal IOPs (N = 4 Eyes/Group); Two-way ANOVA with Bonferroni post-hoc analysis. Figures show datapoints over mean ± SEM, *P < 0.05, ** P < 0.01, *** P < 0.001, **** P < 0.0001

Discussion

This study establishes a mechanistic framework that integrates biochemical and biomechanical risk factors of POAG and highlights the pivotal role of TRPV4, a polymodal Ca2+-permeable channel, as a key regulator of TM contractility and ocular hypertension. Our central finding is that the glaucoma-inducing cytokine TGFβ2 amplifies TRPV4 expression and activity, which in turn drives tonic increases in TRPV4 activation and TM contractility that are required to maintain elevated IOP. These observations position the pressure-activated channel as a molecular linchpin that links mechanical stress to neurodegeneration resulting from the obstruction of the primary outflow pathway. Our confirmation of the linkage between TGFβ signaling and TRPV4 in TM links POAG pathophysiology to fibrotic remodeling seen in other tissues (68, 69). Considering that current glaucoma treatments target secondary outflow mechanisms or incur side effects (such as hyperemia) (70, 71), the IOP lowering achieved through TRPV4 inhibition and gene knockdown promises a novel therapeutic avenue to mitigate ocular injury.

Glaucoma is a multifactorial disease with etiology that reflects the convergence of risk factors that include IOP and TGFβ2: epidemiological data correlate the incidence of POAG with the amplitude of IOP and [TGFβ2]AH (22, 72), while chronic increases of either [TGFβ2]i or IOP promote fibrotic remodeling of TM/SC and augment the flow resistance of the conventional pathway (17, 24, 25). TGFβ2-induced facility suppression has been historically attributed to changes in composition, crosslinking and amount of ECM (25, 26, 73, 74), activation of Hippo signaling and Rho kinase- (Rho/ROCK) mediated contractility (19, 57) and altered expression of genes encoding mitogen-activated protein kinase (MAPK), immune response, oxidative stress, and/or ECM pathways (75–77). Our discovery reveals bidirectional interplay between biomechanical and biochemical mechanisms: TGFβ2 impacts the expression and function of TM mechanosensors and vice versa, TRPV4 is required for TGFβ2-induced contractility: TGFβ2 (i) induced time-dependent upregulation of TRPV4 mRNA and amplified TRPV4-mediated calcium signaling, while (ii) TRPV4 was required to mediate TGFβ2 -induced TM hypercontractility and maintain chronic OHT in TGFβ2-treated mouse eyes. Microinjection of the selective antagonist HC-06 accordingly reduced IOP in LV- TGFβ2-treated eyes to baseline with hypotension persisting for ∼4 days and reversing to pre-injection OHT by day 7. The TRPV4-dependence of TGFβ2-induced OHT and contractility was corroborated in Trpv4-/- mice and in vitro using TGFβ2-treated 3D hydrogel constructs. In vivo, pharmacological inhibition (∼100% reduction in OHT, transient) outperformed gene knockdown (∼50% reduction in OHT, stable), potentially due to compensatory mechanosensory mechanisms in Trpv4-null animals (55).

We’ve previously shown that TM TRPV4 is activated by physiological (5 – 25 mm Hg) pressure steps (39, 60) and (1 – 12%) strains (37, 41), which trigger downstream outflow-relevant signaling through Rho kinase, F-actin, tyrosine phosphorylation of FAK, paxillin and vinculin, lipid remodeling, and ECM release mechanisms (37, 41, 52). The present study extends those observations by revealing bidirectional effects of TRPV4 modulation on TM contractility (the agonist GSK101 induced, and the antagonist HC-06 suppressed, contractility in a 3D biomimetic model) and by establishing TRPV4 as an obligatory effector of OHT under physiological (circadian rhythmicity) and pathological conditions. These findings resolve conflicting reports of hypotensive vs. hypertensive effects of TRPV4 modulation (7, 36, 37, 39, 41, 52, 53, 55). Although TRPV4 activity has been suggested to lower IOP via phosphoinositide signaling in primary cilia (36), TM-resident endothelial nitric oxide synthase (eNOS) (7), release of polyunsaturated fatty acids (PUFAs) (53) and/or signaling downstream from Piezo1 mechanosensing (54), these mechanisms are challenged by evidence that TRPV4-regulated Ca2+ influx persists in TM cells with ablated primary cilia (37), eNOS expression in TM cells is vanishingly low (78–80), PUFAs such as arachidonic acid and EETs stimulate rather than inhibit TRPV4 (37) and TRPV4 signaling in TM cells is unaffected by Piezo1 inhibition and knockdown (39). Instead, the suppression of outflow facility by Piezo1 inhibitors applied under in vitro and in vivo conditions (39, 81) suggests that Piezo1 may oppose the hypertensive functions of TRPV4.

The TRPV4-dependence of increased outflow resistance is indicated by multiple lines of evidence. In biomimetic human TM scaffolds that support flow devoid of ciliary body or SC influences, TRPV4 inhibition enhanced, and activation suppressed, the facility (37). We accordingly demonstrate that the agonist (GSK101) induces contractility while the antagonist (HC-06) mitigates TGFβ2-induced hypercontractility and lowers IOP in TGFβ2-overexpressing eyes. Taken together, these results support a model wherein TRPV4-mediated pressure transduction is augmented by TGFβ2 to drive hypercontractility and fibrosis via Ca2+-and Rho-dependent stress fiber formation and reinforcement of focal ECM contacts (41, 82, 83) (Figure 7). TRPV4 is a thermosensitive ion channel with Q10 of ∼10 and optimal activation at physiological temperature of ∼34 - 37°C (84, 85). TGFβ2 -induced contractility of TM-populated hydrogels thus reflects both the severalfold transcriptional and functional amplification of TRPV4-mediated signaling (Figures 1-3) and constitutive temperature-facilitated activation of TRPV4 channels that is unmasked by the absence of hypercontractility in samples treated with HC-06. The modest contractility observed in HC-06-treated cells may reflect residual contributions from Piezo1, TRPC, and/or TREK-1 channels (37, 60, 86, 87). Parallels from studies conducted on heart, lung, liver, skin and articular cartilage similarly found that TRPV4 contributes to the progression of fibrosis induced by the cognate TGFβ1 (68, 88–91) and implicated the channel in bladder (92), heart (93, 94), and blood vessel (95) contractility. Indeed, conditional Trpv4 ablation from smooth muscle cells lowered blood pressure (96, 97), an effect not dissimilar from IOP lowering in Mgp:Trpv4 cKO mice (Figure 5).

TGFβ2-TRPV4 interactions in TM remodeling and ocular hypertension.

Chronic exposure to TGFβ2 induces upregulation of functional TRPV4 channels alongside the autoinhibitory canonical modulator SMAD7. TRPV4-mediated Ca2+ influx, canonical, and non-canonical TGFβ2 signaling stimulate the Rho/ROCK pathway to augment cytoskeletal contractility, and stimulate ECM release. Actomyosin contractility promotes outflow resistance and drives OHT and underpins a vicious feedforward TRPV4-dependent loop that maintains ocular hypertension. Created with BioRender.com/umbwthh.

TGFβ2-induced upregulation of FN1, SNAIL1, and CTGF transcripts (Figure 1) accords with RNA profiling studies which catalogued the cytokine’s role in transdifferentiation of TM cells into contractile myofibroblasts (25, 75, 98–103) whereas the decreased expression of TGFΒR2 and increased abundance of SMAD7 mRNA indicate concurrent activation of autoinhibitory mechanisms associated with canonical TGFβ-family signaling (104). The-3-fold increase in TRPV4 transcription and responsiveness to GSK101, observed at POAG-relevant TGFβ2 concentrations (In AH ∼0.2-3.2 ng/ml; 20) showed a time course that mirrored facility reduction in human eyes treated with exogenous cytokine (105). A single 5 ng/ml TGFβ2 dose was sufficient to double the amplitude of the GSK101-evoked current and alter its rectification (Figure 3), with imaging experiments confirming robust and reproducible increases in Ca2+ signals across the 5/5 studied strains after 5 days of treatment (Figure 2). The effects of TGFβ2 on ITRPV4, membrane protein levels and [Ca2+]GSK accord with increased expression of the TRPV4 gene. Precedents from other cell types (e.g., fibroblasts) suggest functional upregulation might involve increased trafficking of TRPV4- PI3Kg complexes and/or β-arrestin 1-dependent ubiquitination (106, 107). The upregulation of TRPV4/Piezo1 transcription by TGFβ2 predicts exaggerated responsiveness to mechanical loading, as reported for chemotherapy (108), neuropathic pain (109, 110), cancer (111), and diabetic neuropathy (112).

An important corollary of these findings is that they extend TRPV4’s role beyond pathology into the IOP homeostasis in healthy animals. TRPV4 inhibitors, ROCK and TM-specific expression of dominant negative scAAV2.dnRhoA constructs lower IOP across diverse OHT models (occlusion of the iridocorneal angle, TGFβ, glucocorticoids and the nocturnal cycle; 55, 62, 113). Physiological (circadian-induced) and pathological OHT were not additive (Figure 6) yet TRPV4 inhibition reversibly lowered nocturnal IOP in both untreated and TGFβ2 eyes to imply convergence at the level of TRPV4-Rho signaling as the final effector of OHT maintenance. Future studies should elucidate the mechanisms that mediate the reversibility of circadian TRPV4 activation. We conjecture potential involvement of the suprachiasmatic nucleus, the hypothalamus–pituitary–adrenal axis (114, 115), nocturnal release of norepinephrine and melatonin (116, 117) and altered TRPV4 modulation by β1 integrins (118), caveolin-1 (52), and cytoskeletal proteins (actin, actin adaptor proteins, microtubules) (119).

This study bridges biomechanical and biochemical paradigms of glaucomatous remodeling by extending the role of TGFβ2 beyond fibrosis to include TRPV4-dependent signaling and actomyosin contraction. We propose that TGFβ2 -induced upregulation of TRPV4 expression shifts the normotensive setpoint maintained by steady-state TRPV4, Piezo1 and TREK-1 activity (39, 60, 81) to heighten the TM sensitivity to mechanical cues, hijack contractile machinery and induce fibrosis that facilitates the pull of stress fibers on the increasingly “stiff” ECM (19). The absence of structural and functional visual phenotypes in TRPV4 KO mice (55, 66, 120) predicts that IOP lowering, suppression of fibrosis and protection of retinal neurons from pressure by small-molecule TRPV4 antagonists can take place without compromising homeostatic IOP regulation (70). The conserved TM TRPV4 expression (37) and TM physiology in mice vs. humans (121, 122) suggest that the findings reported might be relevant in a clinical context.

Methods

Animals

C57BL/6J mice were from JAX laboratories (Bar Harbor, ME), Trpv4-/- (Trpv4tm1.1Ldtk Tg(KRT14-cre/ERT)20Efu/0; MGI:5544606) mice were a gift from Wolfgang Liedtke (Duke University) (64, 65). The animals were maintained in a pathogen-free facility with a 12-hour light/dark cycle and ad libitum access to food and water, at a temperature of ∼22-23°C. Mice were 2-6 months in age prior to LV-injection; data from both male and female sexed animals were included in this study.

Human TM Culture

De-identified postmortem eyes from donors with no history of glaucoma (pTM cells) were procured from Utah Lions Eye Bank with written informed consent of the donor’s families. TM cells were isolated from juxtacanalicular and corneoscleral regions as previously described (37, 39), in accordance with consensus characterization recommendations (123). pTM cells were cultured in Trabecular Meshwork Cell Medium (TMCM; Sciencell) in Collagen-I (Corning) coated culture flasks and glass coverslips at 37°C in a humidified atmosphere with 5% CO2. Fresh media was supplied every 2-3 days. Serum free (SF) media was mixed as needed by excluding fetal bovine serum (FBS, Sciencell) from the TMCM. A list of all pTM strains used is available in Table 1; all cells were used between passages 2-4. Cell lines were chosen based on availability at the time of experiments.

For contractility experiments pTM cells were isolated from healthy donor corneal rims discarded after transplant surgery, as previously described (19, 57, 124), and cultured according to established protocols (123, 125). Three pTM cell strains isolated from healthy donors and validated with dexamethasone-induced myocilin expression were used for contractility experiments. pTM cells were cultured in low-glucose Dulbecco’s Modified Eagle’s Medium (DMEM; Gibco; Thermo Fisher Scientific) containing 10% fetal bovine serum (FBS; Atlanta Biologicals) and 1% penicillin/streptomycin/glutamine (PSG; Gibco) and maintained at 37°C in a humidified atmosphere with 5% CO2. Fresh media was supplied every 2-3 days.

The experiments were conducted according to the tenets of the Declaration of Helsinki for the use of human tissue.

Reagents

The TRPV4 antagonist HC-067047 (HC-06) was purchased from Millipore-Sigma or Cayman Biotech and dissolved in DMSO at 20mM. The TRPV4 agonist GSK1016790A (GSK101; Cayman Biotech) was dissolved in DMSO at 1mM. Aliquots were diluted into working concentrations (10-25 nM, GSK101; 5-100 µM, HC-06). Recombinant human TGFβ2 protein was ordered from R&D Systems and reconstituted in sterile 4 mM HCl with 0.1% BSA at 20 ug/mL.

Quantitative Real-Time PCR

Gene-specific primers were used to detect the expression of target genes, as described (126). Total RNA was isolated using the Arcturus PicoPure RNA isolation kit (Thermofisher Scientific). cDNA was generated from total RNA using qScript XLT cDNA Supermix (Quanta Biosciences). SYBR Green based real-time PCR was performed with 2X GREEN Master Mix (Apex Bioresearch Products). Gapdh was used as an endogenous control to normalize fluorescence signals. Gene expression relative to GAPDH was measured using the comparative CT method (2-[ΔCT(gene)- ΔCT(GAPDH)]). All genes were assessed in 4-8 individual samples taken from 3-7 different pTM strains. The primer sequences, expected product length, and gene accession are given in Table 2.

Western Blot

3 SF- or TGFβ2-treated samples were pelleted and pooled together from 3 different pTM samples within the same condition. To separate membrane proteins from heavier cellular debris the pooled cell pellets were homogenized in a hypotonic lysis buffer (20mM TRIS-HCl, 3mM MgCl2, 10mM NaCl, 10mM PMSF, 0.5 mM DTT, 20 mM NaF, 2 mM NaV, 0.5 µg/mL leupeptin) before centrifuging at 300x g for 5 minutes (4 °C). The resulting supernatant was removed and centrifuged again at >12,500 rpm for 30 minutes to pellet membrane proteins, which were then resuspended in RIPA Buffer (Santa Cruz). Proteins were separated on a 10% SDS-PAGE gel and transferred to polyvinylidene difluoride membranes (Bio-Rad). Membranes were blocked with 5% skim milk/2% BSA in TBST and incubated at 4 °C overnight with a primary antibody against TRPV4 (1:250, Alomone Labs #ACC-034) or rabbit antibody against β-tubulin (1:2000, Abcam #EPR1330). Appropriate secondary antibodies conjugated to HRP were used to visualize protein expression on an iBright CL750 imaging system (Thermo Fisher Scientific). β-Tubulin expression was used to standardize protein levels between samples.

Calcium Imaging

pTM cells were seeded onto Collagen-I (Corning) coated coverslips and cultured in TMCM media (ScienCell) as described (39, 41). The cells were serum starved for 24 hours followed by serum-free TMCM with or without TGFβ2 (1 or 5 ng/mL) for 24 hours or five days. The cells were loaded with 10 µM of the ratiometric indicator Fura-2 AM (Kd at RT = 225 nM (Invitrogen/ThermoFisher) for 30-60 minutes. Coverslips were placed in a RC-26G chamber platform (Warner Instrument Corp) and perfused with external saline (pH 7.4) (in mM): 80 NaCl, 4.7 KCl, 1.2 MgCl2, 10 D-Glucose, 19.1 HEPES sodium salt, 2 CaCl2 and osmolality adjusted to 300 mOsm using D-mannitol. External solutions were delivered via a manually controlled gravity-fed eight-line manifold system, with perfusion speed kept constant to minimize changes in shear. Epifluorescence imaging was performed using an inverted Nikon Ti microscope with a 40x 1.3 N.A. oil objective and Nikon Elements AR software. 340 nm and 380 nm excitation were delivered by a high-intensity 150W Xenon arc lamp (Lambda DG-4; Sutter Instruments), high pass-filtered at 510 nm and detected with a 12-bit Delta Evolve camera (Photometrics/Teledyne). GSK101 (10 nM) evoked Δ[Ca2+]i was assessed as ΔR/R (dividing the difference between the peak GSK-evoked F340/F380 signal during stimulation and baseline F340/F380 signal by the baseline F340/F380 signal). Every data point represents a separate experimental day and pTM cell strain, each with 3-5 control and 3-5 TGFβ2-treated slides tested on the same day. TGFβ2 datapoints represent the average GSK101 evoked ΔR/R across all TGFβ2 cells as a % of the average ΔR/R of control cells from the same cell strain on the same day.

Collagen hydrogel contraction assay

Rat tail collagen type I (Corning, Thermo Fisher Scientific) was prepared at a concentration of 1.5 mg/ml according to the manufacturer’s instructions. Five hundred microliters of the hydrogel solution were pipetted into 24-well culture plates. Upon complete collagen polymerization, pTM cells were seeded at 1.5 × 105 cells/well atop the hydrogels and cultured in DMEM + 10% FBS + 1% PSG for 48 hours to facilitate even cell spreading. Next, constructs were cultured in serum-free DMEM + 1% PSG supplemented with: i) control (vehicle: 0.008 mM HCl + 0.0004% BSA; 0.025% DMSO), ii) TGFβ2 (5 ng/ml; R&D Systems), or iii and iv) TGFβ2 + HC067047 (5 µM in DMSO) for 36 hours before carefully releasing the hydrogels from the walls using a sterile 10 µl pipette tip to facilitate contraction. The next morning, fresh serum-free DMEM + 1% PSG supplemented with 0.0025% DMSO (= vehicle) was added to groups i-iii; group iv received serum-free DMEM + 1% PSG supplemented with GSK1016790A (25 nM in DMSO). Plates were longitudinally imaged at 600 dpi resolution with a CanoScan LiDE 300 flatbed scanner (Canon USA) at 0, 15, 30, 60, and 120 minutes. Hydrogel construct size was quantified using FIJI software (National Institutes of Health) (127).

Electrophysiology

Borosilicate patch-clamp pipettes (WPI) were pulled using a P-2000 horizontal micropipette puller (Sutter Instruments), with a resistance of 6-8 MΩ. The internal solution contained (mM): 125 K-gluconate, 10 KCl, 1.5 MgCl2, 10 HEPES, 10 EGTA, pH 7.4. Patch clamp data was acquired with a Multiclamp 700B amplifier, pClamp 10.6 software and Digidata 1440A interface (Molecular Devices), sampled at 5kHz and analyzed with Clampfit 10.7. Current-voltage relationships were assessed using Vm steps from -100 to + 100 mV against a holding potential of -30 mV. Current density was measured as the average current during GSK101 exposure subtracted by the average current from the same cell during baseline perfusion.

IOP Measurements

A TonoLab rebound tonometer (Colonial Medical Supply) was used to measure IOP of awake mice between 12-2 P.M. IOP was determined from the mean of 10-20 tonometer readings. Nocturnal measurements were conducted between 9-10 P.M. in awake animals or under 2.5% isoflurane delivered by a Somnosuite isoflurane vaporizer (Kent Scientific). After animals recovered from intracameral HC-06/PBS injections, IOP was measured daily. IOP was measured every day for 4-5 consecutive days to confirm a stable return to baseline. IOP data for individual cohorts was binned into weeks of experimental time to group values for analysis.

Lentiviral Transduction

Lentiviral stock for TGFβ2 (C226,228S) was purchased from VectorBuilder Inc. (VB170816-1094fnw, pLV[Exp]-CMV> {hTGFB2[NM_003238.3](C226,228S)}) (23). Scrambled control lentivirus was purchased from SignaGen Laboratories (LM-CMV-Null-Puro). Mice were anesthetized with an intraperitoneal IP injection of ketamine/xylazine (90 mg/10 mg/ kg body weight), followed by eyedrops containing 0.5% proparacaine hydrochloride and 1% tropicamide ophthalmic solution to numb the eyes and dilate the pupils. Anesthetized mice were secured to allow stereotaxic injection of lentivirus. Intravitreal injections were conducted by creating a guide hole with a 30-gauge needle 1-2 mm equatorial of the cornea-scleral border followed by insertion of a 12° beveled 33-gauge Hamilton syringe (Hamilton Company) secured to a stereotaxic rig (World Precision Instruments) used to insert the needle 2-3 mm into the eye. Each eye was injected with a 2uL bolus of lentivirus diluted to 1×106 TU/µL over the course of one minute, before the needle was quickly drawn and the pilot hole treated with erythromycin ophthalmic ointment USP (Bausch & Lomb). The efficiency of LV-TGFβ2 OHT induction in WT animals was close to 100%. No differences in observable health post-injection were detected between wild type and Trpv4-/- animals or LV-Ctrl and LV-TGFβ2 injected animals.

Intracameral Microinjections

Mice were anesthetized and treated with eyedrops as above, before being placed on an isothermal heating pad. HC-06 (100 µM) or PBS with DMSO (0.5%) as a vehicle were injected into the anterior chamber using a blunt tip Hamilton syringe (Hamilton Company) through a guide hole made using a 30-gauge needle. At the end of each injection a small air bubble was introduced to seal the cornea and minimize fluid outflow. 0.5% Erythromycin ophthalmic ointment USP (Bausch & Lomb) was applied to the eye after the procedure. Intracameral injections were not associated with observable inflammation, corneal opacity or behavioral changes. For the nocturnal IOP experiments in Figure 6, both eyes of two animals were injected with PBS while two were injected with HC-06. When OHT was stably reestablished a week post-injection, the treatment groups were switched, and experiments repeated resulting in four eyes/treatment group for Figure 6 E-F .

Statistical Analysis

GraphPad Prism 9 was used for statistical analysis. Means are plotted ± SEM unless otherwise noted. One-sample t-tests were used to determine whether TGFβ2 treated groups were significantly different than untreated control groups, while one-way ANOVA or two-way ANOVA along with Tukey or Bonferroni’s multiple comparisons test were used to compare multiple groups.

Study Approval

The animal experimental protocols were conducted in accordance with the NIH Guide for the Care and Use of Laboratory Animals and the ARVO Statement for the Use of Animals in Ophthalmic and Vision Research and were approved by the Institutional Animal Care and Use Committee at the University of Utah.